### 2016-03-04 2016/PAVT/main_eng.tex
Move bibliography to main_single.tex main_single.tex and main_eng.tex are ready for sending to journal

\documentclass[a4paper]{article}
\usepackage{fontspec}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[a4paper]{geometry}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\setmainfont{CMU Serif}
\setsansfont{CMU Sans Serif}

\bibliographystyle{unsrt}
\fontsize{10}{12pt}\selectfont

\begin{document}

\title{Implementation of the parallel asynchronous global search algorithm on Intel Xeon Phi\
\footnote{
The paper was supported by the Russian Science Foundation, project No 15-11-30022 “Global optimization, supercomputing  computations, and applications”}}
\author{K.A. Barkalov, I.G. Lebedev, V.V. Sovrasov, A.V. Sysoyev}
\date{}
\maketitle

\begin{abstract}
The paper considers parallel algorithm for solving multiextremal optimization problems. The issues of implementation of the algorithm on state-of-the-art computing systems using Intel Xeon Phi coprocessor are examined. Two approaches for algorithm parallelization, which take into account information about laboriousness of the objective function computing, are considered. Speed up of the algorithm using Xeon Phi compared to the algorithm using CPU only is experimentally confirmed. Computational experiments are carried out on Lobachevsky supercomputer.
\par
\textit{Keywords:}global optimization, multiextremal functions, dimension reduction, parallel computing, Intel Xeon Phi.
\end{abstract}
\nocite{*}

\begin{thebibliography}{10}

\bibitem{evtushenko}
Evtushenko Yu.G.
\newblock {\em Metody resheniya ekstremal'nykh zadach i ikh primenenie v
  sistemakh optimizatsii [Methods of Solving Extremal Problems and Their
  Application in Optimization Systems]}.
\newblock Moscow: Science, 1982. 432 p.

\bibitem{pinter}
Pinter J.
\newblock {\em Global optimization in action (Continuous and Lipschitz
  optimization: algorithms, imple-mentations and applications)}.
\newblock Kluwer Academic Publishers, 1996, 507 p.

\bibitem{sergKvaDiag}
Kvasov~D.E. Sergeyev~Ya.D.
\newblock {\em Diagonal'nye metody global'noy optimizatsii [Diagonal global
  optimization methods]}.
\newblock Moscow: Fizmatlit, 2008. 352 p.

\bibitem{strSergOptBook}
Sergeyev~Ya.D. Strongin~R.G.
\newblock {\em Global Optimization with Non-convex Constraints. Sequential and
  Parallel Algorithms}.
\newblock Kluwer Academic Publishers, 2000, 704p.

\bibitem{strGergrParOptBook}
Strongin R.G. Gergel V.P. Grishagin V.A.~Barkalov K.A.
\newblock {\em Parallel'nye vychisleniya v zadachakh global'noy optimizatsii
  [Parallel computing in global optimization problems]}.
\newblock Moscow: MSU Publishing, 2013. 280p.

\bibitem{sergKvaDiaPaper}
Sergeyev Ya.D.~Kvasov D.E.
\newblock Global search based on efficient diagonal partitions and a set of
  lipschitz constants.
\newblock {\em SIAM Journal on Optimization}, 16(3):910–937, 2006.

\bibitem{zilinsk}
Žilinskas J.
\newblock Branch and bound with simplicial partitions for global optimization.
\newblock {\em Mathematical Modelling and Analysis}, 13(1):145–159, 2008.

\bibitem{gorodGrishOptBook}
Gorodetskiy S.Yu.~Grishagin V.A.
\newblock {\em Nelineynoe programmirovanie i mnogoekstremal'naya optimizatsiya
  [Nonlinear pro-gramming and multiextremal optimization]}.
\newblock N. Novgorod: UNN Publishing, 2007.

\bibitem{examinArtcle}
Sysoyev A.V. Barkalov K.A. Gergel V.P.~Lebedev I.G.
\newblock Mpi-realizatsiya blochnoy mnogoshagovoy skhemy parallel'nogo
  resheniya zadach global'noy optimizatsii [mpi-implementation of the nested
  block scheme for parallel solving of global optimization problems ].
\newblock In {\em Superkomp'yuternye dni v Rossii: Trudy mezhdunarodnoy
  konferentsii (28-29 sen-tyabrya 2015 g., g. Moskva) [Russian supercomputing
  days: proceedings of the interna-tional conference (Moscow, September 28-29,
  2015)]}, pages 411--419. Moscow: MSU Publishing, 2015.

\bibitem{gklsPaper}
Gaviano M. Lera D. Kvasov D.E.~Sergeyev Ya.D.
\newblock Software for generation of classes of test functions with known local
  and global minima for global optimization.
\newblock {\em ACM Transactions on Mathematical Software}, 29:469--480, 2003.

\bibitem{jonesLipOpt}
Jones D.R. Perttunen C.D.~Stuckman B.E.
\newblock Lipschitzian optimization without the lipschitz constant.
\newblock {\em Journal of Optimization Theory and Applications},
  79(1):157–181, 1993.

\bibitem{gablonskyDirect}
Kelley~C.T. Gablonsky~J.M.
\newblock A locally-biased form of the direct algorithm.
\newblock {\em Journal of Global Optimization}, 21(1):27--37, 2001.

\end{thebibliography}

\end{document}