### 2016-08-03 2016/GOM/paper.tex
Add initial version of GOM paper

% gOMSguide.tex
% v5.0 released July 2015

\documentclass{gOMS2e}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables

\theoremstyle{plain}% Theorem-like structures
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\theoremstyle{remark}
\newtheorem{remark}{Remark}

\begin{document}

%\jvol{00} \jnum{00} \jyear{2015} \jmonth{July}

%\articletype{GUIDE}

\title{GLOBALIZER: A NOVEL SUPERCOMPUTER SOFTWARE SYSTEM FOR SOLVING TIME-CONSUMING GLOBAL OPTIMIZATION PROBLEMS}

\author{
\name{V. P. Gergel\textsuperscript{a}$^{\ast}$\thanks{$^\ast$Corresponding author. Email: latex.helpdesk@tandf.co.uk}
and K. A. Barkalov\textsuperscript{b} and A. V. Sysoev}
\affil{\textsuperscript{a}Taylor \& Francis, 4 Park Square, Milton Park, Abingdon, UK;
\textsuperscript{b}Institut f\"{u}r Informatik, Albert-Ludwigs-Universit\"{a}t, Freiburg, Germany}
\received{v5.0 released July 2015}
}

\maketitle

\begin{abstract}
In this paper, we describe the Globalizer software system for solving global optimization
problems. The system is designed to maximize the use of computational potential by modern
high performance computer systems in order to solve the most time-consuming optimization
problems. Highly parallel computations are facilitated using various distinctive computational
schemes: processing several optimization iterations simultaneously, reducing multidimensional
optimization problems using multiple Peano curve evolvents, multi-stage computing based on
nested block reduction schemes. This novelty leverages supercomputer system capabilities
with shared and distributed memory and with large numbers of processors to efficiently
solve global optimization problems.
\end{abstract}

\begin{keywords}
global optimization, information-statistical theory, parallel computing, high-performance
computer systems, supercomputer technologies
\end{keywords}

\section{Introduction}
Global (or multiextremal) optimization problems are among the most complex problems
in both theory and practice for optimal decision making. In these kinds of problems,
the criterion to be optimized has several local optima within the search domain,
which have different values. The existence of several local optima essentially makes
finding the global optimum difficult, since it requires examining the whole feasible
search domain. The volume of computations for solving global optimization problems
can increase exponentially with increasing number of varied parameters.
\par
In addition, global optimization problem statements are used, as a rule, in the most
complex decision making situations, for example, when a computer-aided design of complex
technologies, products, and systems is conducted. In such problems, the efficiency
criteria are nonlinear, the search domains may be non-contiguous, and, most importantly,
the computational complexity of the functions which the optimized criteria and constraints
are based on can be quite essential.
\par
These global optimization problem features impose special requirements on the quality of
the optimization methods and on the software to implement them. The global optimization
methods should be highly efficient, and the software systems should be developed on a good
professional basis. In general, global optimization problems can be solved using modern
supercomputer systems at a reasonable time and cost by only employing parallel
global optimization algorithms.
\par
The general state of the art in the field of global optimization has been presented in a
number of key monographs (Törn, A., Žilinskas, A., 1989), (Horst, R., Tuy, H., 1990), (Zhigljavsky, A. A., 1991), (Pintér, J. D., 1996), (Strongin, R. G., Sergeyev, Y. D., 2000), (Locatelli, M., Schoen, F., 2013), (Floudas, C. A., Pardalos, M. P., 2016), etc.
The development of optimization methods that use high-performance computer systems to
solve time-consuming global optimization problems is an area receiving extensive
attention – see, for example, (Censor, Y., Zenios, S. A., 1998), (Strongin, R. G., Sergeyev, Y. D., 2000), (Ciegis, R., Henty, D., Kågström, B., Zilinskas, J., 2009), (Luque, G., Alba, E., 2011), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
\par
The theoretical results obtained provide efficient solutions to many applied global
optimization problems in various fields of scientific and technological applications – see,
for example, (Floudas, C. A., Pardalos, M. P., 1996), (Pintér, J. D., 1996), (Luque, G., Alba, E., 2011), (Fasano, G., Pintér, J. D., (2013), (Locatelli, M., Schoen, F., 2013), (Pardalos, M. P., Zhigljavsky, A. A., Zilinskas, J., 2016), etc.
\par
At the same time, the practical implementation of these algorithms for multiextremal
optimization within the framework of industrial software systems is quite limited.
In many cases, software implementations of global optimization algorithms are experimental in
nature and are just used by the developers themselves to obtain results from computational
experiments required for scientific publication. To a large extent, this situation
originates from high development costs for professional software systems that can be
used by numerous users. In addition, global optimization problems could rarely be
solved in an automatic mode because of their complexity. As a rule, the user should
actively control the global search process which implies an adequate level of
qualification in the field of optimization (particularly, the user should know and
understand global optimization methods well).
\par
On the market for global optimization software, one can select from the following systems:
\begin{itemize}
\item LGO (Lipschitz Global Optimization) is designed to solve global optimization
problems for which the criteria and constraints satisfy the Lipschitz condition (see, for example, (Pintér, J. D., 1996)).
The system is a commercial product based on diagonal extensions of one-dimensional
multiextremal optimization algorithms.
\item GlobSol (Kearfott, R. B., 2009) is oriented towards solving global optimization problems as well as
systems of nonlinear equations. The system includes interval methods based on the branch and
bound method. There are some extensions of the system for parallel computations, and it is available to use for free.
\item LINDO (Lin, Y., Schrage.L., 2009) is features by a wide spectrum of problem solving mehtods which
it can be used for – these include linear, integer, stochastic, nonlinear, and global
optimization problems. The ability to interact with the Microsoft Excel software
environment is a key feature of the system. The system is widely used in practical
applications and is available to use for free.
\item IOSO (Indirect Optimization on the basis of Self-Organization) is oriented
toward solving of a wide class of the extremal problems including global optimization
problems (see, for example, (Egorov, I. N., Kretinin, G. V., Leshchenko, I. A., Kuptzov, S. V., 2002)). The system is widely used to
solve applied problems in various fields. There are versions of the system for parallel
computational systems. The system is a commercial product, but is available for trial use.
\item MATLAB Global Optimization Toolkit (see, for example, (Venkataraman, P., 2009)) includes a wide spectrum
of methods for solving the global optimization problems, including multistart methods,
global pattern search, simulated annealing methods, etc. The library is compatible to the
TOMLAB system (see, for example, (Holmström, K., Edvall, M. M., 2004)), which is an additional extension the widely-used MATLAB.
It is also worth noting that similar libraries for solving global optimization problems are
available for MathCAD, Mathematica, and Maple systems as well.
\item BARON (Branch-And-Reduce Optimization Navigator) is designed to solve continuous
integer programming and global optimization problems using the branch and bound method (Sahinidis, N. V., 1996).
BARON is included in the GAMS (General Algebraic Modeling System) system used extensively – see (Bussieck, M. R., Meeraus, A., 2004).
\item Global Optimization Library in R is a large collection of optimization methods
implemented in the R language (see, for example, (Mullen, K. M., 2014)). Among these methods, are stochastic and deterministic global optimization algorithms,
the branch and bound method, etc.
\end{itemize}
\par
The list provided above is certainly not exhaustive – additional information on software
systems for a wider spectrum of optimization problems can be obtained, for example,
in (Rios, L. M., Sahinidis, N. V., 2013), (Mongeau, M.,  Karsenty, H., Rouzé, V., Hiriart-Urruty, J. B., 2000), (Pintér, J. D., 2009), etc.
Nevertheless, even from such a short list the following conclusions can be drawn (see also Liberti, L., 2006):
\begin{itemize}
\item  the collection of available global optimization software systems for practical use is insufficient,
\item the availability of numerous methods through these systems allows complex
optimization problems to be solved in a number of cases, however, it requires a rather
high level of user knowledge and understanding in the field of global optimization,
\item the use of the parallel computing to increase the efficiency in solving complex
time-consuming problems is limited, therefore, the computational potential of modern
supercomputer systems is very poorly utilized.
\end{itemize}
\par
In this paper, a novel Globalizer software system is considered. The development of the
system was conducted based on the information-statistical theory of multiextremal
optimization aimed at developing efficient parallel algorithms for global search – see, for example, (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A., 2013).
The advantage of the Globalizer is that the system is designed to solve time-consuming
multiextremal optimization problems. In order to obtain global optimized solutions
within a reasonable time and cost, the system efficiently uses modern high-performance computer systems.
\par
This paper is further structured as follows. In Section \ref{sec:problem}, a statement of the multidimensional
global optimization problem is presented, and an approach to reducing these to one-dimensional
optimization problems is described. In Section \ref{sec:parallel}, parallel computation schemes are presented,
and parallel optimization methods are described. In Section \ref{sec:arch}, the Globalizer architecture
is examined. In Section \ref{sec:experiments}, the results are presented from computational experiments
that confirm the system’s high level of efficiency. Finally, Section \ref{sec:concl} presents the
conclusions and some ideas for future research.

\section{Multidimensional Global Optimization Problems and Dimension Reduction}
\label{sec:problem}
In this paper, the core class\footnote{In general, the Globalizer can be applied for solving
multicriterial multiextremal multidimensional optimization problems with nonlinear constraints.} of
optimization problems which can be solved using the
Globalizer is examined. This involves multidimensional global optimization problems
without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\} \\
  D=\{y\in \mathbf{R}^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}
\end{array}
\end{equation}
where the objective function (y) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||*||\) denotes the norm in \(\mathbf{R}^N\) space.
\par
Let us further assume that the minimized function \(\varphi(y)\) is defined as a
computational procedure, according to which the value \(\varphi(y)\) can be calculated
for any value of vector \(y\in D\) (let us further call the process of obtaining a
value for the minimized function a trial). As a rule, this procedure is computational-costly i.e.
the overall costs of solving optimization problem (\ref{eq:task}) are determined,
first of all, by the number of trials executed. It is also worth noting the essence of the
assumption on satisfying the Lipschitz condition, since one can construct an estimate
of the global minimum based on a finite number of computed values from the optimized function in this case only.
\par
As has been previously shown by many researchers, finding numerical estimates of globally
optimal extrema implies constructing coverage of the search domain \(D\). As a result,
the computational costs of solving global optimization problems are already very high
even for a small number of varied parameters (the dimensionality of the problem).
A notable reduction in the volume of computations can be achieved when the coverages of the
search domain being obtained are non-uniform, i.e. the series of trial points is dense
only in terms of its nearness to the sought-after globally optimized variants.
The generation of such non-uniform coverages could only be provided in an adaptive
way when the selection of the next trial points is determined by the search information
(the preceding trial points and the values of the minimized function at these points)
obtained in the course of computation. This necessary condition considerably
complicates the computational schemes for global optimization methods, since it
implies a complex analysis of a large amount of multidimensional search information.
As a result, many optimization algorithms use, to some extent, various methods of dimensional reduction.
\par
Within the framework of the information-statistical approach, Peano curves (or evolvents)
\(y(x)\) mapping the interval \([0, 1]\) onto an \(N\)-dimensional hypercube \(D\) are
unambiguously used for dimensional reduction (see, for example, Strongin, R. G., Sergeyev, Y. D., 2000).
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}
\end{equation}
\par
The dimensional reduction scheme considered above superimposes a multidimensional
problem with a minimized Lipschitz function to a one-dimensional problem with a
corresponding minimized function to satisfy the uniform Hölder condition (see Strongin, R. G., 1978, Strongin, R. G., Sergeyev, Y. D., 2000) i.e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
\end{equation}
where the constant \(H\) is defined by the relationship \(H=4L\sqrt{N}\), \(L\) is the
Lipschitz constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem (\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are presented in (Strongin, R. G., Sergeyev, Y. D., 2000).
As an illustration, an approximation of a Peano curve for the third level of density is
shown in Figure \ref{fig:peanoC}. The curve shown in Figure \ref{fig:peanoC} demonstrates the winding order of a
two-dimensional domain; the precision of the Peano curve approximation is determined by the
level of density used in the construction.

\begin{figure}[h!]
    \centering
		\includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
		\caption{A Peano curve approximation for the third level of}
    \label{fig:peanoC}
\end{figure}

\par
The computational scheme obtained as a result of the dimensional reduction consists of the following (see Figure \ref{fig:peanoCUsage}):
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional function \(\varphi(y(x)\),
  \item After determining the next trial point \(x\), a multidimensional image \(у\) in the mapping \(y(x)\) is calculated,
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the multidimensional point \(у\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

\begin{figure}[h!]
    \centering
		\includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
		\caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x)\)}
    \label{fig:peanoCUsage}
\end{figure}

\section{The Parallelization Approach and Global Optimization Algorithms}
\label{sec:parallel}
Let us consider the parallelization methods used widely in the theoretical and
practical application of parallel computing within the context of global optimization problems:
\begin{itemize}
  \item The distribution of the search domain \(D\) among the available computing units
  (data parallelization scheme). In the case of optimization problems, this approach is
  insufficient, since in organizing such computations the subdomain, which contain the
  sought global minimum, will be processed by only one processor, and, therefore,
  all of the remaining processors would perform the excess computations.
  \item The parallelization of the computational schemes for optimization algorithms
  (task parallelization scheme). This approach is also insufficient since the direct
  computational costs for executing optimization algorithms are relatively low
  (the majority of computations in the course of a global search are represented by
  calculations of the optimized function values due to the initial assumption of
  considerable computational costs for such calculations).
  \item The parallelization of the computations executed in order to obtain values
  for the optimized function. This approach can be applied since the most computationally
  costly part of the global search process will be parallelized. However, this method is not
  featured by the generality (the development of parallelization methods is to be performed
  every time from scratch while solving each particular optimization problem).
\end{itemize}
\par
Within the framework of information-statistical theory, a general approach to
parallelization computations when solving global optimization problems has been
proposed --- \textit{the parallelism of computations is provided by means of simultaneously
computing the values of the minimized function \(\varphi(y)\) at several different
points within the search domain \(D\)} – see, for example, (Strongin, R. G., Sergeyev, Y. D., 2000), (Strongin, R. G., K. A., Gergel, V. P., Grishagin, V. A., Barkalov, K. A.,  2013).
This approach provides parallelizaion for the most costly part of computations in the global search process.
\par
The global optimization algorithms implemented in Globalizer will be described step-by-step below.
In Subsection \ref{subsec:corepar}, the core sequential multidimensional algorithm of global search (MAGS) will be presented.
In Subsection \ref{subsec:sharedpar}, a parallel generalization of the MAGS algorithm for parallel computations on
computer systems with shared memory will be described. In Subsection \ref{subsec:distribpar}, the scheme
for parallel computations by multiprocessor systems with distributed memory will be provided.

\subsection{Core Multidimensional Generalized Algorithm of Global Search}
\label{subsec:corepar}
The information-statistical theory of global search formulated in (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000)
has served as a basis for the development of a large number of efficient multiextremal
optimization methods --- see, for example, (Gergel, V. P., 1996, 1997), (Gergel, V. P., Strongin, R. G., 2003, 2005), (Grishagin, V. A., Strongin, R. G., 1984), (Sergeyev Y. D., 1995, 1999), (Sergeyev, Y. D., Grishagin V. A., 1994, 2001), (Sergeyev, Y. D., Strongin, R. G., Lera, D., 2013), (Barkalov K. A., Gergel V. P., 2014), etc.
\par
Multidimensional Algorithm of Global Search (MAGS) established the basis for the
methods applied in Globalizer. The general computational scheme of MAGS can be
presented as follows --- see (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000).
\par
Let us introduce a simpler notation for the problem being solved on a computing node
\begin{equation}
\label{eq:oneDimTask}
f(x) = \varphi(y(x)):x\in [0;1]\}
\end{equation}
\par
The initial iteration of the algorithm is performed at an arbitrary point \(x^1(0,1)\).
Let us assume further \(k\), \(k>1\), iterations of a global search are to be completed.
The selection of the trial point \(k+1\) for the next iteration is performed according to the following rules.
\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}

\subsection{Parallel Computations for Systems with Distributed Memory}
\label{subsec:distribpar}

\section{Globalizer System Architecture}
\label{sec:arch}
\section{Results of Numerical Experiments}
\label{sec:experiments}
\subsection{Parallel Computations Using Intel Xeon Phi processors}
\subsection{Parallel Computations Using General Purpose Graphic Processors}

\section{Conclusions}
\label{sec:concl}
In this paper, the Globalizer global optimization software system was examined for
implementing a general scheme for the parallel solution of globally-optimized decision
making using a wide spectrum of computer systems: with shared and distributed memory,
NVIDIA graphic processors and Intel Xeon Phi coprocessors. This approach enables the
efficient use of modern supercomputer systems to solve the most computational costly
multiextremal optimization problems. The computational experiments have confirmed the
prospects for this proposed approach.

\section*{Acknowledgements}
This research was supported by the Russian Science Foundation, project No 16-11-10150
``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance.''

\end{document}