### 2017-04-11 2017/EUROGEN/eurogen.tex
eurogen: add a placeholder for paper

% EUROGEN 2015 - Template for the extended abstract
%defining format issues
%**********************

%Document Settings___________________________________________________________________

\documentclass[10pt,twocolumn,a4paper]{article}
\special{papersize=210mm,297mm}

%\makeatletter
%\renewcommand\section{\@startsection{section}{1}{\z@}%
%                                  {-3.5ex \@plus -1ex \@minus -.2ex}%
%                                  {2.3ex \@plus.2ex}%
%                                  {\normalfont\normalsize\bfseries}}
%\makeatother

%\makeatletter
%\renewcommand\subsection{\@startsection{subsection}{1}{\z@}%
%                                  {-3.5ex \@plus -1ex \@minus -.2ex}%
%                                  {2.3ex \@plus.2ex}%
%                                  {\normalfont\normalsize\em}}
%\makeatother

%________________________________________________________________________________

%Miscellanneous__________________________________________________________________

\sloppy % better line breaks
\newcommand{\authorname}{\textbf}
\newcommand{\authorgroup}{\emph}
\newcommand{\authoraddress}{\emph}
\newcommand{\authormail}{\emph}
%________________________________________________________________________________

%Packages________________________________________________________________________

%\usepackage[latin1]{inputenc}
\usepackage[utf8]{inputenx}
\usepackage[compact]{titlesec}
%\usepackage[pdftex]{graphicx}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,mathrsfs}
\usepackage{fancyhdr}
\usepackage{calc}
\usepackage{titling}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage[superscript]{cite}
\usepackage{titlesec}
\usepackage[none]{hyphenat}
%________________________________________________________________________________
\titleformat*{\section}{\normalfont\normalsize\bfseries}
\titleformat*{\subsection}{\normalfont\normalsize\em}
\voffset = -1cm
\textheight = 23.7cm
\hoffset = -0.5cm
\textwidth = 17cm
\setlength\parindent{0.5cm}
%Headers_________________________________________________________________________

\pagestyle{fancy}
\fancyhead[L]{EUROGEN 2017}\fancyhead[R]{September 13-15, 2017, Madrid, Spain}
\fancyheadoffset[LE,RO]{\marginparsep + \marginparwidth}
\setlength{\topmargin}{0pt}
%________________________________________________________________________________

%Maketitle Conf__________________________________________________________________
\pretitle{\begin{center}\Large}
\posttitle{\end{center}}

\preauthor{\begin{center}\normalsize \lineskip 0.5em\begin{tabular}[t]{c}}
\postauthor{\end{tabular}\par\end{center}}

\predate{\begin{center}\tiny}
\postdate{\par\end{center}}
%________________________________________________________________________________
%--------------------------------------------------------------------------------

%Article data entry_____________________________________________________________

\title{\vspace{-5ex}\textbf{Instructions to Authors for the Preparation of Extended Abstract}}
\author{
  \authorname{First I. Author*}\\
  \authorgroup{International Center for Numerical Methods in Engineering (CIMNE)}\\
  \authoraddress{Universidad Politécnica de Cataluña,Campus Norte UPC, 08034 Barcelona, Spain}\\
  \authormail{Email: cimne@cimne.upc.edu}\\\\
  %--------------------------
  \authorname{Second B. Author}\\
  \authorgroup{Spanish Association for Numerical Methods in Engineering (SEMNI)}\\
  \authoraddress{Edificio C1, Campus Norte UPC, Gran Capitán s/n, 08034 Barcelona, Spain}\\
  \authormail{Email:semni@cimne.upc.edu}
  %}
}
\date{}

%_______________________________________________________________________________

%Article Body___________________________________________________________________

\begin{document}

%Abstract goes here--------------------------------------------------------------

\twocolumn[

\maketitle

\thispagestyle{fancy} %Insert header in title page

\begin{center}
\line(1,0){465}
\end{center}

\begin{@twocolumnfalse}

\textbf{Summary}\\

The purpose of this document is to provide an example of how authors must format their extended abstracts. The summary should be a very brief description of the work, which should enable an easy indexing. It should not exceed 200 words in length. It must be single spaced, even justified across the full width of the page.\\

\emph{\textbf{Keywords:} list up to 8 (eight) words or group of words in order of priority, separated by commas, which will be used for indexing purposes.}
\begin{center}
\line(1,0){465}
\end{center}

\end{@twocolumnfalse}

]

%--------------------------------------------------------------------------------

\section{Introduction}

Extended abstracts must be delivered through the web submission page, completely “camera-ready” (postscript file or PDF).

\section{General specifications}

The extended abstract must be written in English within a printing box of 17cmx23.7cm, centered in the page. The paper including figures, tables and reference mast have a minimum length of 4 pages and must not exceed 8 pages. Any processing systems can be used (Tex, Latex, Word….).

\section{Title, Authors, Affiliation}

\subsection{Title}

The title should be written centered, in 14pt, boldface Roman. It should be single spaced if the title is more than one line long.

\subsection{Author}

The author's name should include first name, middle initial and surname. It should be written centered, (as shown in the first page of these instructions) in 10pt boldface Roman. The presenting author should be marked with an asterisk.

\subsection{Affiliation}

Author's affiliation should be written centered, (as shown in the first page of these instructions) in 10pt italic Roman.

\section{Text}

The normal text should be written single-spaced, justified, using 10pt (Times New) Roman in two columns.

The first line of each paragraph must be indented 0.5cm (except the first paragraph after a heading). There is not inter-paragraph spacing.

\section{Format of references}

References should be quoted in the text by superscript numbers\cite{Baeck:92, Deb:2002}, and grouped together at the end of the paper in numerical order as shown in these instructions.

The entire list of references at the end of the paper should not be longer than 1/8 of the paper (i.e., 1 full column if the paper is 4 pages long, and 2 columns if the paper is 8 pages long).

\bibliography{bibliography}{}
\bibliographystyle{nature}

\end{document}
__________________________________________________________________________