### 2017-11-15 2018/PAVT_Toropov/PAVT2018_MAM.tex
Text in Russian

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}
\usepackage{amssymb,amsmath}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{url}

\usepackage[T2A]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage[english, russian]{babel}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Parallel Multipoint Approximation Method for Large Scale Optimization Problems 
\thanks{This study was supported by the Russian Science Foundation, project No. 16-11-10150.}}

% a short form should be given in case it is too long for the running head
\titlerunning{Parallel Multipoint Approximation Method}

\author{ Victor P. Gergel%(ORCID 0000-0002-4013-2329)
$^1$ \and
Konstantin A. Barkalov%(ORCID 0000-0001-5273-2471)
$^1$ \and
Evgeny A. Kozinov%(ORCID 0000-0001-6776-0096)
$^1$ \and
Vassili V. Toropov%(ORCID 0000-0002-7017-5055)
$^{1,2}$
}
%
\authorrunning{V.P. Gergel \and K.A. Barkalov \and E.A.Kozinov \and V.V. Toropov}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{$^1$Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia\\
\email{\{victor.gergel, konstantin.barkalov, evgeny.kozinov\}@itmm.unn.ru}\\
$^2$Queen Mary University of London, London, UK\\
\email{v.v.toropov@qmul.ac.uk}\\
}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem''
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}
TBD

\keywords{design optimization, multidisciplinary optimization, multipoint approximation method, parallel computing.}
\end{abstract}

\section{Introduction}

\Russian
В статье рассматривается the multipoint approximation method (MAM) и его применение к large scale optimization problems.

In problems with a large (in the order of hundreds) number of design variables, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) proved to be efficient, e.g. in turbomachinery applications \cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010} . This method is an iterative optimization technique based on mid-range approximations built in trust regions. A trust region is a sub-domain of the design space in which a set of design points, produced according to a small-scale design of experiments (DoE), are evaluated. These and a subset of previously evaluated design points are used to build metamodels of the objective and constraint functions that are considered to be valid within a current trust region. The trust region will then translate and change size as optimization progresses. The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values \cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures \cite{ToropovMarkineHolden1999}. The mid-range approximations used in the trust regions, as originally suggested in \cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e. nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure. This was enhanced by the use of gradient-assisted metamodels \cite{ToropovFilatov1993}, use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming \cite{ToropovAlvarez1998}. One of the recent developments \cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e. a two stage approximation building process that is conceptually similar to the original one used in \cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.

The Moving Least Squares Method (MLSM) was proposed in \cite{LancasterSalkauskas1981} for smoothing and interpolation of scattered data and later used in the mesh-free form of the FEM \cite{Liszka1984}. As suggested in \cite{ChoiYounYang2001}, it can be used as a technique for metamodelling and used in MDO frameworks. The MLSM is a weighted least squares method where the weights depend on the Euclidean distance from a sample point to where the surrogate model is to be evaluated. The weight value for a certain sample point decays as the distance increases. Describing the weight decay with a Gaussian function tends to be the most useful option even though many others have been evaluated in \cite{ToropovSchrammSahaiJones2005}. As demonstrated in \cite{PolynkinToropov2010}, the cross-validated MLSM can be used both for design variable screening and for surrogate modelling. In order to create an efficient MDO framework for problems with disparate discipline attributes \cite{OllarToropovJones2014} extended the optimization approach of MAM to the use of local DOEs and MLS approximations built in different subspaces of the total design variable space corresponding to the individual disciplines. The subspaces are finally combined into the total design variable space in which the resulting MDO problem is solved.

\Russian
This paper presents a Parallel Multipoint Approximation Method that makes it capable of handling problems with the number of design variables in the order of thousands. Параллельный вариант алгоритма был разработан с целью минимизировать время, затрачиваемое на работу непосредственно оптимизационной части алгоритма. При этом предполагается, что процесс вычисления значений целевой функции уже распараллелен.

\section{The Multipoint Approximation Method}
\label{sec:MAM}

It is useful to start with a brief description of MAM. A typical formulation of a constrained optimization problem that MAM works with is as follows:
\begin{equation}
  \label{eq:problem}
  \begin{array}{c}
  \min\limits_{a_i \le x_i \le b_i}F_0(x) \\
  s.t.\; F_j(x) \le 1,\; j=1,\dots ,M,
  \end{array}
\end{equation}
where $x$ is the vector of design variables, $a$ and $b$ are the lower and upper bounds for the design variables, respectively, $F_0(x)$ is the objective function, and $F_j(x)$ are the constraints. The numbers of design variables and constraints are $n$ and $M$, respectively. MAM attempts to solve this problem by using approximations of the objective function and constraints in a series of trust regions. The trust region strategy seeks to zoom in on the region where the constrained minimum is achieved. It aims at finding a trust region that is sufficiently small for the approximations to be of sufficiently good quality to improve the design, and that contains the point of the constrained minimum as its interior point. The main loop of the MAM is organized as follows.

Algorithm (MAM).
\begin{enumerate}
\item Initialization: choose a starting point $ x^0$ and initial trust region $[ a^0,  b^0]$ such that $ x^0 \in [ a^0,  b^0]$.
\item At the $k$-th iteration the current approximation to the constrained minimum is $ x^k$, the current trust region is $[ a^k,  b^k] \subset [ a^0,  b^0]$.
  \begin{enumerate}[label=(\alph*)]
    \item Design of Experiments (DoE): a set of points $ x_k^i \in [ a^k,  b^k]$ is chosen to be used for approximation building. Responses are evaluated at the DoE points and approximations are built using the obtained values. Currently, the pool of approximation methods available in MAM consists of a metamodel assemblies \cite{PolynkinToropov2012} and the moving least-squares metamodels \cite{LancasterSalkauskas1981,Liszka1984,ChoiYounYang2001,ToropovSchrammSahaiJones2005}. Other metamodel types could be used as well.

    Denote the approximate objective function and constraints by $\widetilde{F}^k_0(x)$ and $\widetilde{F}^k_j(x)$, respectively.
    \item The original optimization problem (\ref{eq:problem}) is replaced by the following problem:
    \begin{equation}
      \label{eq:problem_approx}
      \begin{array}{c}
      \min\limits_{a_i^k \le x_i ^k\le b_i}\widetilde{F}^k_0(x) \\
      s.t.\; \widetilde{F}^k_j( x) \le 1,\; j=1,\dots ,M,
      \end{array}
    \end{equation}
    The approximate problem (\ref{eq:problem_approx}) is solved using Sequential Quadratic Programming (SQP) and the solution of this problem determines the center of the next trust region.
    \item The size of the next trust region is determined depending on the quality of approximations at the previous iteration, on the history of the points $ x^k$, and on the size of the current trust region \cite{KeulenToropovMarkine1996}.
    \item The termination criterion is checked (it is a part of the trust region strategy and depends on the position of the point $ x^{k+1}$ in the current trust region, the size of the current trust region and the quality of approximations). If the termination criterion is satisfied, the algorithm proceeds to step 3. Otherwise, it returns to step 2.
  \end{enumerate}
  \item Optimization terminates. The obtained approximation to the solution of the problem (\ref{eq:problem}) is $ x^{k+1}$.
\end{enumerate}

The selection of the approximations $\widetilde{F}^k_j(x),\; j=0, \dots ,M,$ is such that their evaluation is inexpensive as compared to the evaluation of the original response functions $F_j(x)$. For example, intrinsically linear functions were successfully used for a variety of design optimization problems in the works \cite{ToropovFilatov1993,KeulenToropov1997}. The approximations are determined by means of the weighted least squares:
\begin{equation}\label{eq:least_sq}
\min \sum_{p=1}^P{w_{pj}\left[ F_j(x_p)- \widetilde{F}^k_j(x_p,a_j) \right]^2}.
\end{equation}
In (\ref{eq:least_sq}) minimization is carried out with respect to the tuning parameters $a_j$; $w_{pj}$ are the weight coefficients, and $P$ is the number of sampling points in Design of Experiments (DoE) which must not be less than the number of parameters in the vector $a_j$.

The weight coefficients $w_{pj}$ strongly influence the difference in the quality of the approximations in different regions of the design variable space. Since in realistic constrained optimization problems the optimum point usually belongs to the boundary of the feasible region, the approximation functions should be more accurate in such domain. Thus, the information at the points located near the boundary of the feasible region is to be treated with greater weights. In a similar manner a larger weight can be allocated to a design with a better objective function, see \cite{ToropovFilatov1993,KeulenToropov1997}.

As optimization steps are carried out, a database with response function values becomes available. In order to achieve good quality approximations in the current trust region, an appropriate selection of DoE points must be made. In this work, DoE points in each trust region is generated randomly. Generally, points located far from the current trust region would not contribute to the improvement of the quality of the resulting approximations in the trust region. For this reason only points located in the neighborhood of the current trust region are taken into account, as depicted in Fig.~\ref{fig:trustregion}.
\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{trustregion.png}
    \caption{Current trust region (smaller box) and its extension (larger box): points outside the larger box are not used for building the approximate functions}
    \label{fig:trustregion}
\end{figure}
A box in the space of design variables, which is approximately 1.5 to 1.8 times larger than the box representing the current trust region, was found by numerical experimentation to be a reasonable choice for the size of the neighborhood.

In this work an approach is used that is based on the assembly of different approximate models $\{\varphi_l\}$ into one metamodel using the following form (note that the indices $j$ and $k$ are suppressed to simplify notation):
\begin{equation}\label{eq:assembly}
\widetilde{F}(x) = \sum_{l=1}^{NF}{b_l\varphi_l(x)},
\end{equation}
where $NF$ is the number of regressors in the model pool $\{\varphi_l\}$ and $b_l$ are corresponding regression coefficients.
The used procedure consists of two subsequent steps. In the first step, the parameters $a_l$ of individual functions (regressors) $\varphi_l$ in (\ref{eq:assembly}) are determined by solving a weighted least squares problem using a specified a DoE of $P$ points:
%\begin{equation}\label{eq:least_sq}
\[
\min \sum_{p=1}^P{w_{p}\left[ F(x_p)- \varphi_l(x_p,a_l) \right]^2},
\]
%\end{equation}
where minimization is carried out with respect to the tuning parameters $a_l$.

In the second step, based on the same DoE and keeping the obtained parameters $a_l$ fixed, a vector $b$ in (\ref{eq:assembly}) is estimated using the following formulation
\[
\min \sum_{p=1}^P{w_{p}\left[ F(x_p)- \widetilde{F}(x_p,b) \right]^2},
\]
that leads to solving a linear system of $NF$ equations with $NF$ unknowns $b_l$, where $NF$ is the number of regressors in the model pool $\{\varphi_l\}$.

The selection of the regressors $\{\varphi_l\}$ is based on the number of sampling points currently located in the trust region. In the mid-range approximation framework, inexpensive approximate models for objective and constraint functions are built using minimum required number of sampling points. The simplest case is a linear function of the tuning parameters $a$:
\[
\varphi(x) = a_0 + \sum_{i=1}^N{a_i x_i}.
\]

This structure can be extended to an \textit{intrinsically linear} function. Such functions are nonlinear, but they can be led to linear ones by simple transformations. The most useful function among them is the multiplicative function
\[
\varphi(x) = a_0 \prod_{i=1}^N{x{_i}^{a_i}}.
\]

Intrinsically linear functions have been successfully used for a variety of design optimization problems. 
The advantage of these approximation functions is that a relatively small number $N+1$ ($N$ is number of design variables) of tuning parameters $a_i$ is to be determined, and the corresponding least squares problem is solved easily. This is the most important feature of such approximations as it allows applying them to large scale optimization problems.

Other intrinsically linear functions may be considered in the model pool, e.g.
\[
\varphi(x) = a_0 + \sum_{i=1}^N{a_i/x_i},
\]
\[
\varphi(x) = a_0 + \sum_{i=1}^N{a_i x_i^2},
\]
\[
\varphi(x) = a_0 + \sum_{i=1}^N{a_i/x_i^2},
\]
\[
\varphi(x) = a_0 + \sum_{i=1}^N{a_i x_i^3},
\]
\[
\varphi(x) = a_0 + \sum_{i=1}^N{a_i/x_i^3}.
\]

As more points are added to the database the approximations may be switched to higher quality models, e.g. a rational model
\begin{equation}\label{eq:rational}
\varphi(x) = \frac{a_1 + a_2x_1 + a_3x_2+...+a_{n+1}x_n}{1 + a_{n+2}x_1 + a_{n+3}x_2+...+a_{2n+1}x_n},
\end{equation}
The coefficients in (\ref{eq:rational}) are determined using a least squares approach which reduces to a nonlinear optimization problem with a constraint on the sign of the denominator (positive or negative). The latter is necessary in order to prevent the denominator from crossing the zero axis within a specified trust region. One may note that this formulation may yield the objective function with many local minima. Currently this problem is resolved using optimization restarts from a specified number of initial guesses randomly generated in a trust region.

Tests results demonstrated that, although the above functions may describe the global behavior rather poorly, such approximations proved to be efficient in the mid-range approximation framework of MAM.

\section{Parallel Multipoint Approximation Method}\label{sec:par_alg}

\Russian
Рассмотрим возможные способы распараллеливания, которые можно применить для рассматриваемой задачи.
Во-первых, можно распараллеливать вычисление функций, описывающих оптимизируемый объект. Данный путь является как очевидным, так и необходимым, т.к. в industrial design optimization problems расчет даже одного значения функции может занимать несколько часов. Однако данный способ является специфичным для каждой конкретной задачи. Вопросы распараллеливания тут решаются на уровне пакетов прикладных программ, в которых выполняется инженерное моделирование (например, Ansys, OpenFOAM, и т.п.).

Во-вторых, можно скорректировать алгоритм с целью параллельного вычисления нескольких значений целевой функции и ограничений в разных точках области поиска. В соответствии с правилами MAM на каждой итерации в текущем trust region формируется P sampling points in design of experiments. Значения функции в них можно вычислять параллельно на разных процессорах (узлах). Сказанное соответствует распараллеливаню шага 2а алгоритма. The number of sampling points generated in each iteration may be set equal to P = k×NP number, where NP is a number of available processors (or nodes) and k ≥ 1. In terms of the time, the latter will be equivalent to NP function evaluations per step. Данный способ был успешно реализован в работе [35] и показал хорошую эффективность, т.к. при числе design variables порядка ста основное время затрачивается на function evaluations, работа самого MAM (в последовательном режиме) вносит незначительные накладные расходы.

Однако когда число переменных становится порядка тысячи, работа последовательной части MAM начинает оказывать ощутимое влияние на общее время решения задачи (в предположении, что время вычисления значений целевой функции не меняется). Так, для рассмотренной в Section 4 задачи, время проведения одной итерации метода увеличилось в более чем в 2000 раз (с 4 до 9000 секунд ) при увеличении числа переменных в задаче с 100 до 1000.

Поэтому в рамках проводимого исследования был реализован третий подход к распараллеливанию алгоритма. Были распараллелены вычислительные правила МАМ, обеспечивающие построение аппроксимации, решение аппроксимированной задачи и выбор следующего trust region (шаги 2b, 2c и 2d алгоритма).

С целью выявления наиболее трудоемких мест алгоритма нами был использован Intel VTune Amplifier XE. Проведенный анализ показал, что наиболее трудоемкими операциями, использованными при реализации MAM, являются матричное умножение и решение СЛАУ при построении аппроксимации, а также решение аппроксимированной задачи методом SQP.

Матричное умножение и решение СЛАУ являются стандартными операциями, реализованными во многих высокопроизводительных библиотеках. Здесь нами были использованы соответствующие параллельные методы из библиотеки Intel MKL. Метод SQP был распараллелен самостоятельно с использованием технологии OpenMP.

\section{Numerical Example}
\label{sec:num_example}

The example considered in this study is a classical engineering optimization problem known as the scalable cantilevered beam \cite{Vanderpllaats2001}. The engineering object to be optimized is shown in Fig. \ref{fig:beam} taken from \cite{Vanderpllaats2001}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=1.0\textwidth]{beam.png}
    \caption{The cantilevered beam}
    \label{fig:beam}
\end{figure}

The design variables are the widths $b_i$ and heights $h_i$ of the segments. The number of segments $N$ can be chosen arbitrary. The total length of the beam is 500 cm, the lengths of the segments are $l_i=500/N$  cm. There are $N$ geometric constraints (the aspect ratios of each block (i.e. heights divided by widths) should not exceed 20) and $N$ constraints on the stress, calculated at the left end of each segment (stresses should not exceed  $\bar{\sigma}=14000\; N/cm^2$). There is also also a constraint on the displacement at the tip, which should not exceed 2.5 cm. The load is $P = 50 000\; N$, the Young's modulus is $E=2\cdot 10^7  N/cm^2$.

The deflection $y_i$ at the right hand of the $i$-th segment is given by the following recursive formulas:
\begin{displaymath}
  \begin{array}{c}
    y_0=y_0'=0, \\
    y_i'=\frac{P\cdot l_i}{E\cdot I_i}\Big[ L+\frac{l_i}{2}-\sum\limits_{j=1}^i l_j\Big]+y_{i-1}', \\
    y_i=\frac{P\cdot l_i^2}{2E\cdot I_i}\Big[L-\sum\limits_{j=1}^i l_j + \frac{2l_i}{3}\Big]+y_{i-1}'l_i+y_{i-1}.
  \end{array}
\end{displaymath}
The moment of inertia of a segment is $I_i=\frac{b_i h_i^3}{12}$ and the bending moment at the left end is $M_i=P[L+l_i- \sum_{j=1}^{i}  l_j ]$. The maximum bending stress in the segment $i$ is then given by the following formula:
\begin{displaymath}
  \sigma_i=\frac{M_i h_i}{2I_i}
\end{displaymath}

We are looking for a design with smallest volume $V = \sum_{i=1}^N b_i h_i l_i$. The widths $b_i$ vary from 1.0 to 10.0 cm and the heights $h_i$ from 5.0 to 100.0 cm. The optimization problem is formulated as follows:
\begin{displaymath}
  \begin{array}{c}
    \min\limits_{ b,  h}V( b,  h) \\
    s.t.\;1.0\le b_i \le 10.0, \\
    5.0 \le h_i \le 100.0, \\
    y_N\le 2/5, \\
    \sigma_i \le \bar{\sigma}=14000, \\
    \frac{h_i}{b_i}\le 20.
  \end{array}
\end{displaymath}

With $N=50$ segments (corresponding to 100 design variables) the SQP solution of this problem is $V = 63704.598\; cm^3$. The optimal values of design variables are given below (the first 50 entries are the widths and the last 50 entries are the heights of the segments):
\setcounter{MaxMatrixCols}{11}
\[
\begin{matrix}%[pos]{spalten}
			b=& 3.246& 3.224& 3.202& 3.179& 3.156& 3.133& 3.109& 3.085& 3.061& 3.036 \\  
			  & 3.011& 2.985& 2.959& 2.933& 2.905& 2.878& 2.850& 2.821& 2.792& 2.762 \\
			  & 2.731& 2.700& 2.668& 2.635& 2.602& 2.567& 2.532& 2.495& 2.458& 2.419 \\  
			  & 2.379& 2.338& 2.295& 2.250& 2.204& 2.156& 2.105& 2.052& 1.996& 1.936 \\  
			  & 1.873& 1.805& 1.732& 1.651& 1.562& 1.462& 1.345& 1.196& 1.023& 1.000 \\ 	
\end{matrix}
\]
\[
\begin{matrix}
	h=& 64.919&64.480&64.033&63.581&63.122&62.656&62.184&61.703&61.216&60.720  \\
		& 60.216&59.703&59.182&58.651&58.110&57.558&56.996&56.423&55.837&55.239  \\
		& 54.628&54.003&53.363&52.707&52.035&51.344&50.635&49.906&49.155&48.381  \\
		& 47.582&46.754&45.897&45.008&44.082&43.116&42.105&41.041&39.919&38.729  \\
		& 37.462&36.102&34.630&33.020&31.240&29.241&26.910&23.919&20.465&14.639  \\
\end{matrix}
\]

MAM obtained the solution $V = 63935.360\; cm^3$, using 1500 function evaluations (as compared to almost 8000 evaluations used by SQP). The number of points in a trust region used to build approximations was 150. The optimal values of design variables obtained by MAM are given below:
\[
\begin{matrix}
	b=&	3.238&3.204&3.202&3.170&3.176&3.128&3.108&3.089&3.057&3.003 \\  
		& 3.027&2.986&2.959&2.952&2.885&2.855&2.864&2.847&2.803&2.762 \\  
		& 2.737&2.722&2.630&2.645&2.590&2.558&2.547&2.480&2.693&2.391 \\ 
		& 2.368&2.310&2.307&2.227&2.176&2.149&2.106&2.016&2.007&1.925 \\  
		& 1.864&1.843&1.758&1.635&1.582&1.934&1.332&1.173&1.026&2.419 \\ 
\end{matrix}
\]
\[
\begin{matrix}
	h=& 64.764&64.083&64.036&63.392&63.518&62.566&62.153&61.772&61.149& 60.054  \\
		& 60.534&59.728&59.182&59.033&57.695&57.105&57.274&56.943&56.057&55.233  \\
		& 54.747&54.433&52.590&52.902&51.802&51.150&50.943&49.597&51.567&47.811  \\
		& 47.373&46.190&46.140&44.543&43.525&42.991&42.119&40.308&40.141&38.506  \\
		& 37.268&36.839&35.153&32.702&31.615&28.304&26.642&23.447&20.536& 9.417

\end{matrix}
\]

The solution obtained by MAM is very close to the reference solution obtained by SQP, except for the last design variable (the height of the last segment), which indicates that the problem is insensitive to this variable near the optimum, making it hard for the metamodels to capture this dependence. Both SQP and MAM solutions are, however, feasible and differ only slightly in the value of the objective function.

\Russian
Далее сравним время работы последовательной и параллельной версий программы при решении large scale problems. Размерность N будем варьировать от 100 до 1000 design variables, что соответствует 50 – 500 сегментам cantilevered beam. The number of points in a trust region used to build approximations was 1.5 N  Таблица 1 отражает время работы параллельного алгоритма в зависимости от числа переменных. В таблице 2 приведено число итераций MAM, а также число вычислений значений функции, потребовавшееся для решения задачи. Для сравнения приведено также число вычислений значений функции, которые бы потребовались для решения исходной (неаппроксимированной) задачи методом SQP. Во всех проведенных экспериментах значения целевой функции в последовательной и параллельной версиях алгоритма совпадают (с точностью до вычислительной погрешности), и не сильно отличаются от решения, полученного SQP. При этом наглядно демонстрируется сокращение числа function evaluations, требующееся для решения задачи с использованием MAM по сравнению с использованием SQP.

%Таблица

Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhny Novgorod. The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The CPU had 8 cores, i.e. a total of 16 physical cores were available in the node. MS Visual Studio 15 and Intel Fortran Compiler were used to implement the algorithm.

\section{Conclusions}

Recent developments in the Multipoint Approximation Method (MAM) made it capable of solving large scale industrial optimization problems. The approach uses metamodels built in the space of design variables. MAM solves initial problem by using approximations of the objective function and constraints.
\Russian
В рамках данного исследования разработана параллельная версия МАМ, ориентированная на сокращение времени работы оптимизационного алгоритма (в предположении, что вычисление значений the objective function and constraints уже распараллелены). Проведенные эксперименты показывают приемлемое ускорение при решении задач большого размера при использовании 16 ядер.

Performance is demonstrated on a benchmark example of structural optimization known as the scalable cantilevered beam.

\bibliography{bibliography}{}
\bibliographystyle{nature}

%\bibitem{Strongin2000}
%Strongin, R.G., Sergeyev, Y.D.: Global Optimization with Non-Convex Constraints. Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000); DOI: 10.1007/978-1-4615-4677-1
%
%\bibitem{Barkalov2002}
%Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)
%
%\bibitem{Gergel2005}
%Gergel, V.P., Strongin, R.G.: Parallel computing for globally optimal decision making on cluster systems. Future Gener. Comput. Syst. 21(5), 673--678 (2005); DOI: 10.1016/j.future.2004.05.007
%
%\bibitem{Strongin2013}
%Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to global optimization exploiting space-filling curves. Springer (2013);  DOI: 10.1007/978-1-4614-8042-6
%
%\bibitem{Barkalov2010}
%Barkalov, K., Ryabov, V., Sidorov, S.: Parallel Scalable Algorithms with Mixed Local-Global Strategy for Global Optimization Problems. In: Hsu, Ch., Malyshkin, V. (Eds.) MTPP 2010, LNCS, vol. 6083, pp. 232--240. Springer, Heidelberg (2010); DOI: 10.1007/978-3-642-14822-4$\_$26
%
%\bibitem{Barkalov2015}
%Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi Coprocessor for Solving Global Optimization Problems. In: Malyshkin, V. (Ed.) PaCT 2015, LNCS, vol. 9251, pp. 307--318. Springer, Heidelberg (2015); DOI: 10.1007/978-3-319-21909-7$\_$31
%
%\bibitem{Barkalov2016}
%Barkalov, K., Gergel, V.: Parallel Global Optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016); DOI: 10.1007/s10898-016-0411-y
%
%\bibitem{Gergel2016}
%Barkalov, K., Gergel, V., Lebedev, I.: Solving Global Optimization Problems on GPU Cluster. In: Simos T.E. (Ed.) ICNAAM 2015, AIP Conference Proceedings, vol. 1738, art. no. 400006 (2016); DOI: 10.1063/1.4952194
%
%\bibitem{Gergel2015}
%Gergel, V., Sidorov, S.: A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems. In: Malyshkin, V. (Ed.) PaCT 2015, LNCS, vol. 9251, pp. 505--515. Springer, Heidelberg (2015); DOI: 10.1007/978-3-319-21909-7$\_$49
%

%\end{thebibliography}

\end{document}