### 2018-06-25 2018/Journal of Supercomputing/JoS_Paper.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex 
%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%
%
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn

%\usepackage[utf8x]{inputenc}
%\usepackage[english,russian]{babel}

\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{multirow}
%
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
 \journalname{The Journal of Supercomputing}
%
\begin{document}

\title{Parallel Computing for Constrained Global Optimization Problems
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-
10150.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Konstantin Barkalov          \and
        Ilya Lebedev %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{K. Barkalov  \at
               Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
              \email{konstantin.barkalov@itmm.unn.ru} 
           \and
           I. Lebedev \at
					Lobachevsky State University of Nizhni Novgorod, Nizhni 
Novgorod, Russia \\
							\email{ilya.lebedev@itmm.unn.ru} 
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This paper considers a parallel algorithm for solving multiextremal problems with non-convex 
constrains. Within the framework of the approach proposed here, the solution of the initial 
multidimensional problem is reduced to solving a set of linked one-dimensional problems in 
parallel. The parallelization scheme does not require a single control process since the solution of 
any one-dimensional problem is a solution of the initial multidimensional one. The non-convex 
constraints are accounted for by a novel index scheme without utilizing the ideas of penalty 
functions. The efficiency of the parallel algorithm was confirmed experimentally by solving 
numerically several hundred multidimensional multiextremal problems with non-convex 
constrains generated randomly.
\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and 
Dimensionality reduction \and Parallel algorithms}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}

The aim of this paper is to examine parallel methods for solving global optimization problems 
with non-convex constraints. The objective function and constraints are assumed to satisfy the 
Lipschitz condition with a priori unknown Lipschitz constants. The analytical form of the 
problem's functions may also be unknown, i.e. they can be set by an algorithm computing their 
values at various points within the search domain (the so-called ``black-box'' functions). It is 
also supposed that even a single computation of the function value can be time-consuming, 
since it requires performing numerical simulation in applied problems.

The numerical algorithms for solving the problems of the above-mentioned class are noted for 
an exponential growth of the computational costs with increasing dimensionality (the so-called 
``curse of dimensionality''). The use of supercomputers expands the field of application of the 
global optimization methods and, however, at the same time, it gives rise to the problem of 
efficient parallelization of the search process.

Parallel algorithms utilizing heuristic approaches and based in one way or another on the idea of 
random search are at present quite popular (see \cite{RefFerreiro,RefZhu,Korosec,Guerrero}). 
However, because of their probabilistic nature, the algorithms of such type guarantee the 
convergence to the global optimum solution in the probabilistic sense only, which makes them 
less attractive than deterministic parallel global optimization algorithms 
\cite{Jones2001,Paulavicius2011,Evtushenko2013,Strongin2000}. As a rule, efficient 
deterministic algorithms are intended to solve unconstrained optimization problems. The 
problems with constraints are solved using the penalty function method that has a number of 
disadvantages. In particular, this method cannot be applied to the problems with the 
\textit{partially defined functions} (i.e. when any constraint is violated, the values of all the 
other functions of the problem remain undefined).

This situation is typical for many optimal design problems, when at violation of some conditions 
of functioning of the modeled object, other characteristics of this one appear to be undefined. 
For example, the modeled object is an electronic device, and its response time is minimized. 
However, the response time is undefined if at given combination of parameters the device is not 
working at all.

%Another example is the optimal control problem described through system of ordinary 
%differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}. It is only 
%possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz 
%matrix, i.e. every eigenvalue of $A$ has a strictly negative real part. Otherwise, the value of the 
%criteria is indeterminate.

In this study, we used a novel approach to the minimization of multiextremal functions with 
non-convex constraints developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002}. The 
approach referred to as the index method of accounting for the constraints is based on 
accounting for each constraint of the problem separately and is not related to the use of the 
penalty functions. The solution of the initial multidimensional problem is reduced (by using the 
Peano-type space-filling curves) to solving an equivalent set of one-dimensional problems, 
which can be solved in parallel in such a way that each of these processes shares the information 
obtained by the other processes. The proposed scheme was implemented on the Lobachevsky 
supercomputer with the use of the distributed memory as well as the shared one.

The main part of the paper has the following structure. Section 2 states the constrained 
optimization problem, reviews the index method and an approach to reducing dimensionality by 
using Peano curves. Section 3 presents a parallel implementation of the index method using a set 
of space-filling curves. Section 4 presents the results of numerical experiments. Section 5 
concludes the paper.

\section{Problem Statement}
\label{sec:2}

A constrained global optimization problem can be formulated as follows
%Let us consider the $N$-dimensional global optimization problem
\begin{equation}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
\end{equation}
\begin{equation}\label{D}
D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.
\end{equation}
The objective function $\varphi(y)$ (hereafter denoted by $g_{m+1}(y)$) and the left-hand 
sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
\[
\left|g_i(y_1)-g_i(y_2)\right|\leq L_i\left\|y_1-y_2\right\|, \;1\leq i\leq m+1, \; y_1,y_2 \in D,\;
\]
with a priori unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal. It is 
assumed that the functions $g_i(y)$ are defined and computable only at the points $y \in D$ 
satisfying the conditions
\begin{equation}\label{g_k}
g_k(y) \leq 0, \; 1 \leq k < i.
\end{equation}

By employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval 
$[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the 
minimum in (\ref{problem}) by solving the one-dimensional problem
\[
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\]
Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are 
considered in \cite{Sergeyev2013}. Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined 
and computable in the subranges 
\[
Q_1=[0,1], \; Q_{i+1}=\left\{x \in Q_i : g_i(y(x)) \leq 0 \right\}, \; 1 \leq i \leq m.
\]

These conditions allows us to introduce a classification of the points $x \in [0,1]$ according to 
the number $\nu (x)$ of the constraints computed at this point. The \textit{index} $\nu(x)$ can 
also be defined by the conditions
\begin{equation}\label{nu}
g_i(y(x)) \leq 0, \; 1 \leq i < \nu, \; g_\nu(y(x))>0,
\end{equation}
where the last inequality is inessential if $\nu=m+1$.

In the dimensionality reduction scheme considered here, a multidimensional problem with 
Lipschitzian functions is juxtaposed with a one-dimensional problem, where the corresponding 
functions satisfy uniform H{\"o}lder condition (see \cite{Sergeyev2013}), i.e.,
\[
\left|g_i(y(x_1))-g_i (y(x_2))\right| \leq H_i \left|x_1-x_2 \right|^{1/N}, \; x_1,x_2\in [0,1], \; 
1\leq i \leq m+1.
\]
Here, $N$ is the dimensionality of the initial multidimensional problem and the coefficients 
$H_i$ are related to the Lipschitz constant $L_i$ of the initial problem as $H_i \leq 2L_i 
\sqrt{N+3}$.

Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm 
will consist of the following sequence of operations:
\begin{itemize}
	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping 
$y(x)$;
	\item Compute the values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from 
(\ref{nu}). 
\end{itemize}
The dyad $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ will be referred to as the \textit{trial 
outcome}.

The scheme of the serial index algorithm is as follows. The first trial is executed at an arbitrary 
internal point $x_1 \in (0,1)$. The selection of the point $x^{k+1}, \; k \geq 1,$ of any next trial 
is determined by the following steps.

Step 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the lower indices in the 
increasing order of the coordinate values, i.e.
\[
0=x_0<x_1<\dots <x_k<x_{k+1}=1,
\]
and juxtapose to them the values $z_i=g_\nu(y(x_i)), \; \nu=\nu(x_i), \; 1 \leq i \leq k,$ computed 
at these points. The points $x_0=0$ and $x_{k+1}=1$ are introduced additionally and the 
values $z_0$ and $z_{k+1}$ are not defined.

Rule 2. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} 
$R(i)$ using some formulae \cite{Barkalov2002}.

Rule 3. Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
\begin{equation}\label{MaxR}
R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation}

Rule 4. Execute the next trial at the internal point of the interval $(x_{t-1},x_t)$, i.e. $x^{k+1} 
\in (x_{t-1},x_t)$.

Step 5. Check termination condition $\left|x_t-x_{t-1}\right|^{1/N}\leq \epsilon$, where $t$ is 
the number of interval with the maximum characteristic and $\epsilon > 0$ is the predefined 
accuracy.

A detailed description of this algorithm and the corresponding theory of convergence are 
presented in \cite{Strongin2000,Sergeyev2001,Barkalov2002}.

The comparison of two different methods of accounting for the constraints -- the classical 
penalty functions method and the index scheme -- is presented in \cite{Barkalov2017}. In 
this paper, the advantage of the index method over the penalty functions one in the number of 
iterations as well as in the number of computations of the constraints and of the objective 
function has been confirmed experimentally. To solve a problem with given accuracy, the index 
method required 3-5 times less iterations (subject to the problem dimensionality and the feasible 
domain fraction). At the same time, the objective function value was computed in average at 
every fifth iteration of the index method only, in difference from the penalty functions method, 
in which the objective function values were computed at every iteration.

\section{Parallel Index Algorithm with a Set of Evolvents}
\label{sec:3}

The reduction of the multidimensional problems to the one-dimensional ones using evolvents 
has such important properties as the continuity and preservation of boundedness of function's 
divided differences. However, a partial loss of information on the nearness of the points in the 
multidimensional space occurs in this case since a point $x \in [0,1]$ has only the left and the 
right neighbors while the corresponding point $y(x) \in R^N$ has the neighbors in $2N$ 
directions. As a result, when using the mappings like a Peano curve, the close images $y', y'' \in 
D$ can correspond to the remote preimages $x', x'' \in [0,1]$. This property results in excess 
computations since several limit points $x', x''$ of the trial sequence generated by the index 
method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional 
space.

One of the possible ways to overcome this disadvantage consists in using the multiple mapping 
$Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$. To construct the set 
$Y^S(x)$, different approaches can be used. 
%For example, in \cite{Strongin2000} a scheme was implemented, according to which each 
%evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent 
%$y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed 
%allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 
%1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in 
%one coordinate.

For example, a scheme of building of the multiple mapping was proposed in \cite{6_Gergel2009}. The building of a set of Peano curves by rotation of the evolvents 
around the coordinate origin is a distinctive feature of this scheme. In the initial non-rotated 
evolvent for close points $y', y''$ in the multidimensional space their preimages $x', x''$ in the 
interval $[0,1]$ can be far away from each other. In the rotated scheme there exists a mapping 
$y^i(x)$ according to which the preimages $x', x''$ will be located closer to each other. The 
maximum number of various rotations of the evolvents mapping the $N$-dimensional 
hypercube onto a one-dimensional interval is $2^N$. The employment of all possible rotations 
might appear to be redundant. In this case, we can select only a part of all rotations. As a 
possible approach, we propose to generate the new evolvents by the rotation of the initial 
evolvent on the angle of $\pm\pi/2$ in each coordinate plane. Taking into account the initial 
evolvent, the total number of such evolvents for mapping the $N$-dimensional domain onto the 
corresponding one-dimensional intervals equals to $N(N-1)+1$. This method for building a set 
of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
Just such an approach to building the set of evolvents was employed in the present paper when 
carrying out the numerical experiments.

With the use of multiple mapping, it is possible to solve the initial problem (\ref{problem}) by 
solving in parallel the problems
\[
\min \left\{\varphi(y^s(x)): x \in [0,1], \; g_i(y^s(x))\leq 0, \; 1 \leq i \leq m\right\},\; 1\leq s \leq S.
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a 
separate processor/node. The trial outcome at the point $x^k$ obtained for the problem being 
solved by particular processor are interpreted as the outcomes of the trials in the remaining 
problems (at the corresponding points $x^{k_1},...,x^{k_S}$). In this approach, a trial at the 
point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists of the following 
sequence of operations.

\begin{enumerate}

	\item Determine the image $y^k=y^s(x^k)$ for the evolvent $y^s(x)$.

	\item Inform the rest of processors about the start of the trial execution at the point 
$y^k$ (\textit{the blocking} of the point $y^k$).

	\item Compute the values $g_1(y^k),...,g_\nu(y^k)$, where the index $\nu \leq m$ is 
determined by the conditions
	\[
		g_i(y^k)\leq 0, \; 1 \leq i < \nu, \; g_\nu(y^k)>0, \; \nu \leq m.
	\]
The trial terminates at the point as soon as the first violation of any constraint occurs. In the case 
when $y^k$ is a feasible one, the trial includes the computation of the values of all problem 
functions. In this situation, the index is set to $\nu=m+1$. The triplet
\[
y^s(x^k), \; \nu=\nu(x^k), \; z^k=g_\nu(y^s(x^k))
\]
is the outcome of the trial at the point $x^k$.

\item Determine the preimages $x^{k_s} \in [0,1]$, $1 \leq s \leq S$, of the point $y^k$ and 
interpret the trial executed at the point $y^k \in D$ as the execution of the trials at the $S$ 
points  $x^{k_1},...,x^{k_S}$ with the same outcomes
\[
 \nu(x^{k_1}) = ... = \nu(x^{k_S}) = \nu(x^k),
\]
\[
 g_\nu(y^1(x^{k_1})) = ... = g_\nu(y^S(x^{k_S})) = z^k.
\]

\item Inform the rest of processors about the trial outcome at the point $y^k$.

\end{enumerate}

The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the 
sequential algorithm (with the exception the method of the trial execution). To organize the 
interactions between the processors, the queues are created on each processor, where the 
processors store the information on the executed iterations in the form of tuples: the processor 
number $s$, the trial point $x^{k_s}$, the index $\nu(x^{k_s})$, and the value $g_\nu(y^s
(x^{k_s} ) )$. Moreover, the index of the blocked point is assumed to be equal to $-1$; the 
function value at this point is undefined.

The proposed parallelization scheme was implemented with the use of MPI and OpenMP 
technologies. A separate MPI-process is created for each of $S$ one-dimensional problems 
being solved, one process per one processor employed. Each process uses $p$ threads for 
solving a one-dimensional problem, one thread per an accessible core.

\section{Results of Numerical Experiments}
\label{sec:4}

A well-known approach to investigating and comparing the multiextremal optimization 
algorithms is based on testing these methods by solving a set of problems, chosen randomly 
from some specially designed class.

In this study, we will use a GKLS generator of the functions of arbitrary dimensionality 
proposed in \cite{Gaviano2003} to produce constrained problems. The scheme for generating 
constrained global optimization problems is proposed in \cite{Gergel2017}. 
The test problems with constraints obtained using this scheme have an important property 
inherent to the applied optimization problems: the constrained global minimum is located at the 
feasible domain boundary whereas the global minimum of the objective function without 
constraints is located outside the feasible domain. At that, one can specify the feasible domain 
volume and the number of constraints active at the minimum point thus complicating or 
simplifying the problem statement.

In the present study, an experimental investigation is carried out of the speedup, which is 
obtained by the use of the index method in combination with the two-level parallelization 
scheme from \cite{Sidorov2015}.

The experiments have been carried out by solving a series of $100$ problems with two 
constraints and one objective function from the \textit{Simple} and \textit{Hard} GKLS 
classes with the dimensionalities $N=4$, $N=5$. The number of the used cluster nodes $S$ 
and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at 
each node have been varied. The problem was considered to be solved, if the algorithm 
generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-
y^*\right\|\leq \delta = 0.03\left\|b-a\right\|$. For the purpose of simulation of the computational 
complexity inherent to applied optimization problems, calculation of the problem functions in all 
performed experiments was made more complex by additional calculations without changing 
the type of function and arrangement of its minima (a series summation of $100$ thousand 
elements).

Computational experiments were carried out on Lobachevsky supercomputer. The node of 
supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The 
CPU had 8 cores, i.e. each node had a total of 16 cores. The average time and the number of 
iterations, which were required to solve the problems of the series at various parallelization 
parameters are given in Tables 1 and 2. Here, \textit{Node(core)} are the numbers of employed  
nodes and cores per node, respectively.

\begin{table}
	\caption{Average time}
	\label{tab:1}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
		\multirow{2}{*}{\textit{Node(core)}} & \multicolumn{2}{c}{ $N=4$ } & & 
\multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1(1)	&	220.5	&	334.8	&	&	1223.6	&	1386.6	\\
1(16)	&	31.3	&	49.1	&	&	211.8	&	547.2	\\
2(1)	&	158.4	&	260	&	&	1052.9	&	1458.1	\\
2(16)	&	22.1	&	35.9	&	&	227.5	&	603	\\
4(1)	&	127.7	&	286.4	&	&	951.3	&	1362.2	\\
4(16)	&	20.9	&	45	&	&	206	&	925.7	\\
8(1)	&	99.3	&	141.8	&	&	700.1	&	897.3	\\
8(16)	&	31	&	77.7	&	&	264.6	&	374	\\
10(1)	&	99.3	&	141.8	&	&	700.1	&	897.3	\\
10(16)	&	31	&	77.7	&	&	264.6	&	374	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Average number of iterations}
	\label{tab:2}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
		\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & 
& \multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1(1)	&	58320	&	84546	&	&	266943	&	287102	\\
1(16)	&	4297	&	6601	&	&	22655	  &	56754	\\
2(1)	&	34791	&	52126	&	&	188465	&	241369	\\
2(16)	&	2029	&	3239	&	&	16689	  &	40763	\\
4(1)	&	22223	&	47771	&	&	135734	&	180489	\\
4(16)	&	1281	&	2483	&	&	9241	  &	35024	\\
8(1)	&	13844	&	18933	&	&	77748	  &	94563	\\
8(16)	&	608	  &	1473	&	&	5820	  &	23033	\\
10(1)	&	13844	&	18933	&	&	77748	  &	94563	\\
10(16)	&	608	&	1473	&	&	5820	  &	23033	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

The results demonstrate the speedup achieved when using the shared memory at a node as well 
as the distributed memory. In this case, the highest time speedup was 10 (when using 64 cores 
on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster 
nodes). The difference in the speedups in time and in the number of iterations can be explained 
by the effect of the overheads of the data transmission between the processes. Note that when 
solving applied optimization problems, the computation of the problem function values even in 
one point is a computationally costly operation. In this case, the data transfer overheads will not 
affect significantly the total computational costs, and the difference between the time speedup 
and the iteration speedup will not be so great.

\section{Conclusions}
\label{sec:5}

In conclusion, note that the majority of the optimization algorithms (except the methods based 
on the random search ideas) implement the purposeful search strategy. This strategy consists in 
excluding many non-prospective cases from further consideration on the base of 
analysis of a small part of variants and concentrating further search in the subset containing the 
best variant.

The complexity of the modeled processes and objects can result in the non-monotonous
characteristics of these ones that causes the inadequacy of local search 
methods for evaluating the optimal variant. In such problems, the global search methods should 
be applied, which provide the purposeful search, for example, by means of accounting for the 
limited variation of the characteristics of the modeled object at limited variations of its 
parameters. The mathematical formulation of this fact can take the forms of the Lipschitz 
condition, uniform H\"older condition, etc. The parallel index algorithm considered in the 
present paper belongs just to such methods capable to solve the complex applied optimization 
problems.

The parallel index algorithm %for solving constrained global optimization problems %considered in this paper 
offers the following possibilities:
\begin{itemize}
	\item to solve the initial problem directly, without the use of the penalty functions (thus 
eliminating the need to select the penalty coefficient and to solve a series of unconstrained 
problems with different penalty coefficients);
	\item to solve the problems where the values of the problem function are not defined 
everywhere (for example, the objective function values are undefined outside the feasible 
domain);
	\item to use the two-level parallelization scheme with the shared and distributed memory 
proposed earlier for the unconstrained optimization methods.
\end{itemize}

The parallel algorithm has demonstrated a speedup with respect to the number of 
processors/cores employed. This was confirmed by the results of solving numerically of several 
hundred test problems using 160 cores of the Lobachevsky supercomputer.

The application of the set of evolvents for the dimensionality reduction in the multicriterial 
optimization problems would be a promising direction of further work. In such problems, the 
construction of Pareto set requires solving even a series of scalar problems. When solving these 
ones in parallel, one can expect a higher speedup than when solving a single scalar problem.

% Non-BibTeX users please use
\begin{thebibliography}{}

\bibitem{RefFerreiro}
Ferreiro, A.M., Garcia, J.A., Lopez-Salas, J.G., Vazquez, C.: An efficient implementation of 
parallel simulated annealing algorithm in GPUs. J. Glob. Optim. 57(3), 863--890 (2013)

\bibitem{RefZhu}
Zhu, W.: Massively parallel differential evolution--pattern search optimization with graphics 
hardware acceleration: an investigation on bound constrained optimization problems. J. Glob. 
Optim. 50(3), 417--437 (2011).

\bibitem{Korosec}
Koro\v sec, P., Vajter\v sic, M.,  \v Silc, J., Kutil, R.: Multi-core implementation of the 
differential ant-stigmergy algorithm for numerical optimization. J. Supercomput. 63(3), 757--
772 (2013)

\bibitem{Guerrero}
Guerrero, G.D., Cecilia, J.M., Llanes, A. et al.: Comparative evaluation of platforms for parallel 
Ant Colony Optimization. J. Supercomput. 69(1) 318--329 (2014)

\bibitem{Jones2001}
Jones, D.R.: The direct global optimization algorithm. In: Floudas, C.A., Pardalos, P.M. (eds.) 
The Encyclopedia of Optimization, Second Edition. pp. 725--735. Springer (2009)

\bibitem{Paulavicius2011}
Paulavi\v cius, R., \v Zilinskas, J., Grothey, A.: Parallel branch and bound for global 
optimization with combination of Lipschitz bounds. Optim. Methods Softw. 26(3), 487--498 
(2011)

\bibitem{Evtushenko2013}
Evtushenko, Y.G., Posypkin, M.A.: A deterministic approach to global box-constrained 
optimization. Optim. Lett. 7(4), 819--829 (2013)

\bibitem{Strongin2000}
Strongin, R.G., Sergeyev, Ya.D.: Global Optimization with Non-Convex Constraints. 
Sequential and Parallel Algorithms. Kluwer Academic Publishers, Dordrecht (2000)

%\bibitem{Balandin2011} Balandin, D.V., Kogan, M.M.: Optimal linear-quadratic control: 
From matrix equations to linear matrix inequalities. Automation and Remote Control. 72(11), 
2276--2284 (2011)

%\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-
control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 
(2017)

\bibitem{Sergeyev2001}
Sergeyev, Y.D., Famularo D., Pugliese P. Index Branch-and-Bound Algorithm for Lipschitz 
univariate global optimization with multiextremal constraints. J. Glob. Optim. 21(3), 317--341 
(2001)

\bibitem{Barkalov2002}
Barkalov, K.A., Strongin, R.G.: A global optimization technique with an adaptive order of 
checking for constraints. Comput. Math. Math. Phys. 42(9), 1289--1300 (2002)

\bibitem{Sergeyev2013}
Sergeyev, Ya.D., Strongin, R.G., Lera, D.: Introduction to Global Optimization Exploiting 
Space-Filling Curves. Springer (2013)

\bibitem{Gaviano2003}
Gaviano, M., Lera, D., Kvasov, D.E., Sergeyev, Ya.D.: Software for generation of classes of 
test functions with known local and global minima for global optimization. ACM Trans. Math. 
Software 29, 469--480 (2003)

\bibitem{Gergel2017}
Gergel, V.: An approach for generating test problems of constrained global optimization. 
LNCS. 10556, 314--319 (2017)

%\bibitem{Barkalov2015} Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi coprocessor 
for solving global optimization problems. LNCS. 9251, 307--318 (2015)

\bibitem{Sidorov2015}
Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of 
computationally intensive multiextremal optimization problems. LNCS. 9251, 505--515 (2015)

\bibitem{Barkalov2017}
Barkalov, K., Lebeled, I.: Comparing Two Approaches for Solving Constrained Global 
Optimization Problems. LNCS. 10556, 301--306 (2017)

\bibitem{6_Gergel2009}
Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem 
solving. Journal of instrument engineering. 52, 25--33 (2009) (In Russian)

%
%\bibitem{RefJ}
%% Format for Journal Reference
%Author, Article title, Journal, Volume, page numbers (year)
%% Format for books
%\bibitem{RefB}
%Author, Book title, page numbers. Publisher, place (year)
% etc
\end{thebibliography}

\end{document}
% end of file template.tex