### 2018-07-30 2018/LeGo/LeGo.tex
final

\documentclass{aip-cp}

\usepackage[numbers]{natbib}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{enumitem}

%\pdfmapfile{+txfonts.map}

%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
%\usepackage[utf8x]{inputenc}
%\usepackage[english,russian]{babel}
%\usepackage{cmap}
%%%%%

% Document starts
\begin{document}

% Title portion

\title{A Flexible Generator of Constrained Global Optimization Test Problems}

\author{Victor Gergel}
\author{Konstantin Barkalov\corref{cor1}}
%\eaddress[url]{http://www.aip.org}
\author{Ilya Lebedev}
%\eaddress{anotherauthor@thisaddress.yyy}
\author{Maria Rachinskaya}
\author{Alexander Sysoyev}

\affil{Lobachevsky State University of Nizhni Novgorod, Gagarin avenue 23, 
Nizhni Novgorod, Russia, 603950}

\corresp[cor1]{Corresponding author: konstantin.barkalov@itmm.unn.ru}

\maketitle

\begin{abstract}
In the present work, further development of an approach to constructing test global 
optimization problems with nonlinear constraints is considered. In the generated problems, the 
location of the global minimum is known. The considered generator is featured by an option of 
specifying desired number of constraints and the fraction of feasible domain relative to the 
whole global search domain. In the updated version of the generator, the options of forming the 
problems with a conditional minimum located at the boundary of the feasible domain and of 
controlling the number of constraints active at the minimizer have been added. The 
demonstration of the developed approach is given on the example of a known index method for 
solving complex multiextremal optimization problems with non-convex constraints.
\end{abstract}

% Head 1
\section{INTRODUCTION}

In the present paper, the methods for generating global optimization test problems with non-convex constraints
\begin{eqnarray}
&\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}, 
\label{i_problem} \\
&D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\} \label{D}
\end{eqnarray}
are considered. The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and 
the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints are supposed to satisfy the 
Lipschitz condition
\[ \left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1, \]
with the Lipschitz constants unknown a priori. The analytical formulae of the problem functions 
may be unknown, i.e. they may be defined by an algorithm for computing the function 
values in the search domain (so called ``black-box''-functions).

%\Russian
%Проблема построения генераторов многоэкстремальных тестовых задач с невыпуклыми ограничениями является актуальной, т.к. стандартный подход к исследованию и сравнению алгоритмов многоэкстремальной оптимизации основан на решении серии тестовых задач.
%Известные генераторы порождают, как правило, задачи без ограничений. Например, способ построения двумерных тестовых функций был предложен Гришагиным (описание доступно в \cite{Grishagin1997,Grishagin2015}). Another generator (\textit{GKLS generator}) for the functions of arbitrary dimensionality has been proposed in \cite{Gaviano2003}.  GKLS-generator был успешно применен при исследовании некоторых методов глобальной оптимизации \cite{Paulavicius2014,Sergeyev2015,Grishagin2018}.

%На основе генератора GKLS был предложен подход к построению задач с ограничениями (Emmental-type GKLS-based test problems, \cite{Sergeyev2017}. В данном генераторе в качестве целевой функции используется модифицированная GKLS-функция, ограничениями являются exterior of a ball с заданными центрами и радиусами. Использование иных целевых функций и/или ограничений недоступно, что является недостатком данного генератора.

The problem of developing the generators of the multiextremal test problems with non-convex 
constraints is relevant since classes of test problems are used usually for investigating and comparing global optimization algorithms.
Well known generators construct the problems without constraints as a rule. For example, a 
method for constructing two-dimensional test functions has been proposed by Grishagin (its 
description can be found in \cite{Grishagin1997,Grishagin2015}). Another generator 
(\textit{GKLS generator}) for the functions of arbitrary dimensionality has been proposed in 
\cite{Gaviano2003}.  GKLS generator has been successfully applied in the investigation of several global 
optimization methods \cite{Paulavicius2014,Sergeyev2015,Grishagin2018}.
Based on the GKLS generator, an approach to constructing the constrained problems has been 
proposed (Emmental-type GKLS-based test problems \cite{Sergeyev2017}). In this generator, 
the modified GKLS-function was used as the objective function, the constraints were the 
exterior of the balls with given centers and radii. The use of other objective functions and/or 
constraints was not allowed that was a disadvantage of this generator.

In the present work, further development of the approach to generating the global optimization 
problems with non-convex constraints proposed originally in \cite{Gergel2017} has been 
conducted.
When generating the test problems, the necessary number of constraints and the desired fraction 
of the feasible domain relative to the whole global search domain can be specified. 
At that, the location of the global minimum is known for the generated problems that essentially simplifies 
evaluating the results of the numerical experiments. 
In addition to the specified options, the following functions have been added in the current version of the generator:

\begin{itemize}
\item The possibility to generate the problems with the constrained minimizer located at the 
boundary of the feasible domain.
\item The possibility to control the number of constraints active at the optimum point.
\end{itemize}

The above capabilities allow emulating the properties of the applied constrained global 
optimization problems more adequately.

\section{TEST PROBLEM CLASSES}

The generator GCGen (Global Constrained optimization problem Generator) which allows to 
generate the test global optimization problems with $m$ constraints has been proposed in \cite{Gergel2017}. In the mentioned paper the rules which allow formulating the constrained global optimization problems have been proposed so that:
\begin{itemize}
	\item one could control the size of feasible domain with respect to the whole domain of 
the parameters' variation;
	\item the global minimizer of the objective function would be known a priori taking into 
account the constraints;
	\item the global minimizer of the objective function without accounting for the 
constraints would be out of the feasible domain;
	\item the value of the gradient at the global solution of the constrained problems is zero.
\end{itemize}

The location of the minimizer at the boundary of the feasible domain is an important 
property featuring the applied constrained optimization problems. 
When generating the problems according to the scheme described in \cite{Gergel2017}, 
the constrained minimum point, in general, will be located inside the feasible domain. Therefore, the 
options of shifting the minimizer to the boundary of the feasible domain as well of 
controlling the number of the active constraints at the global solution have been added in the new version 
of the problem generator.

The shifting of the minimizer to the boundary of the feasible domain is performed in 
the following way.
\begin{enumerate}
\item A problem with an arbitrary location of the constrained minimum point $\overline{y}$ is 
generated;
\item From the point $\overline{y}$, a coordinate search of a feasible point $y^*$, for which at 
least one constraint is active (i.e. there is such index $j, 1 \leq j\leq m$ that $\left|g_j(y^*)\right| 
\leq \delta$) is performed with given step $h$. 
\item A rough estimate of $y^*$ obtained is refined by a local method. 
\item The coordinate transformation transferring the point $\overline{y}$ to the point $y^*$ is 
performed. This way, the objective function reaches its minimum at the boundary of feasible 
domain. 
\end{enumerate}

Note that since the search of the point $y^*$ is performed in each coordinate separately, the 
computation costs would increase linearly with increasing the dimensionality of the problem.
%\Russian 
At that, no more than $2N\left\|b-a\right\|/h$ computations of the objective function and 
constraints values will be required.

In the case, when it is necessary to generate a problem with a number of active 
constraints at the minimizer not less than $S$, the rules are modified as follows.
\begin{enumerate}
\item A problem with an arbitrary location of a constrained minimizer $\overline{y}$ is 
generated;
\item From the point $\overline{y}$, a coordinate search of a feasible point $y^*$, for which at 
least one constraint is active (i.e. there is such index $j, 1 \leq j\leq m$, that $\left|g_j(y^*)\right| 
\leq \delta$) is performed with a predefined step $h$.
\item A rough estimate $y^*$ obtained is refined by a local method. 
\item The number $K$ of the constraints active at the point $y^*$ is determined.
\item If $K<S$, $S-K$ constraints are selected from the set of the inactive ones, and these 
ones are set to be the active by adding a positive parameter $q$ to the right hand side.
\item The coordinate transformation transferring the point $\overline{y}$ to the point $y^*$ is 
performed. This way, the objective function reaches its minimum at the boundary of the feasible 
domain with the number of active constraints not less than $S$. 
\end{enumerate}

\section{NUMERICAL RESULTS}

As an illustration, the level lines of the objective functions and the zero-level lines of three 
constraints for the problem constructed on the base of functions from \cite{Grishagin2015} 
are shown in Fig.~\ref{fig:VAG}. The feasible domains are highlighted by green. 
Figures~\ref{fig:VAG} (a,b,c,d) correspond to different methods of generating the test problems.

\begin{enumerate}[label=(\alph*)]
	\item The problem has been generated with the constrained optimizer located inside the 
feasible domain.
\item A modification of the problem from item (a): unconstrained optimizer is located outside 
the feasible domain.
\item A modification of the problem from item (b): constrained optimizer is located at the 
boundary of the feasible domain, the number of active constraints is not predefined.
\item A modification of the problem from item (c): the number of active constraints is defined 
to be two.
\end{enumerate}

%\Russian
Note that the volume fraction of the feasible domain for the first three problems will be equal to 
$\Delta = 0.5$, and will be no more than $\Delta = 0.5$ for the last problem (because the 
volume of feasible domain can decrease when generating the problems with controlled number 
of active constraints).

\begin{figure}[ht]
  \centering
  \begin{tabular}[b]{cc}
    \includegraphics[width=.42\linewidth]{1.png} & \includegraphics[width=.42\linewidth]{2.png}\\
    \small (a) & \small (b) \\
    \includegraphics[width=.42\linewidth]{3.png} & \includegraphics[width=.42\linewidth]{4.png}\\
    \small (c) & \small (d) 
  \end{tabular}
  \caption{The test problems generated by GCGen}
	\label{fig:VAG}
\end{figure}

Figures~\ref{fig:VAG} (a,b,c,d) also shows the points of 425, 381, 289 and 188 trials, 
correspondingly, performed by the \textit{index method} for solving constrained global 
optimization problems until the required accuracy $\epsilon=10^{-2}$ was achieved. The 
conditional optimizer is shown as a red point and the best estimation of the optimizer is shown 
as a blue point.
The index method has been developed in 
\cite{Strongin2000,Sergeyev2001,Barkalov2002}. The approach is based on a separate 
accounting for each constraint of the problem and is not related to the use of the penalty 
functions.

%\Russian
The index objective function utilized in the method essentially differs from the classical penalty one 
 since the computing of the initial objective function value outside the feasible 
domain is not required in the index method. Moreover, it is not required to compute the values 
of all constraints outside the feasible domain as well but until the occurrence of the first 
violated one only whereas in the classical penalty problems all constraints along with the 
objective function are computed at every point.
This allows the index method: 
(i) accounting for the information on each constraint separately and (ii) solving the problems, in 
which the function values may be undefined out of the feasible domain. 
%It should be noted, that the index method can be efficiently parallelized for accelerators 
%\cite{Barkalov2015,Barkalov2016}.

\section {Conclusion}

This paper considers the method for generating the global optimization test problems with non-convex constraints, which allows:
\begin{itemize}
	\item controlling the size of the feasible domain with respect to the whole domain of the 
parameters' variation;
	\item knowing a priori the conditional global minimizer of the objective 
function;
	\item generating the unconditional global minimizer of the objective function out of the 
feasible domain;
	\item generating the problems with a conditional minimizer located at the boundary 
of the feasible domain;
	\item controlling the number of constraints active at the optimum point.
\end{itemize}

The developed approach allows generating any number of test global optimization problems 
with non-convex constraints for performing computational experiments in order to 
obtain a reliable evaluation of the efficiency of the developed optimization algorithms.

% Acknowledgement
\section{ACKNOWLEDGMENTS}
This research was supported by the Russian Science Foundation, project No 16-11-10150.

% References

%\nocite{*}
\bibliographystyle{aipnum-cp}%
\bibliography{LeGo}%

\end{document}