### 2019-02-28 2019/NUMTA_Barkalov/Abstract/Barkalov.tex

% Latex-template for one-page abstract submissions.
% NUMTA2019 - Numerical Computations: Theory and Algorithms
% 15-21 June 2019, Le Castella - Isola Capo Rizzuto, Crotone (Italy)

\documentclass[oribibl]{llncs}

% Please, do not use your own macros/redefinitions.

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}
\cleardoublepage

\title{Title of Your Presentation}
\author{Roman Strongin, Konstantin Barkalov, Semen Bevzuk}

%%% Please note that the first author is supposed to be the speaker.

\institute{University of Nizhni Novgorod, Nizhni Novgorod, Russia\\
\email{president@unn.ru, barkalov@vmk.unn.ru}}

\maketitle

\thispagestyle{fancy}

\textbf{Keywords.} Global optimization; keyword2; keyword3.

 \vspace*{0.5cm}

%-----------------------------------------------------------------

The goal of the Conference is to create a multidisciplinary round
table for an open discussion on numerical modeling nature by using
traditional and emerging computational paradigms. The Conference
will discuss all aspects of numerical computations and modeling from
foundations and philosophy to advanced numerical techniques. New
technological challenges and fundamental ideas from theoretical
computer science, linguistic, logic, set theory, and philosophy will
meet requirements and new fresh applications from physics,
chemistry, biology, and economy.

Researchers from both theoretical and applied sciences are very
welcome to submit their papers and to use this excellent possibility
to exchange ideas with leading scientists from different research
fields. Papers discussing new computational paradigms, relations
with foundations of mathematics, and their impact on natural
sciences are particularly solicited. A special attention will be
also dedicated to numerical optimization and different issues
related to theory and practice of the usage of infinities and
infinitesimals in numerical computations.

\textbf{Acknowledgements.}

This research was supported by the the Russian Science Foundation, project No.\,16-11-10150.

\vspace{0.5cm}

\begin{thebibliography}{4}

\bibitem{Book} Strongin R.G., Sergeyev Ya.D. (2000) \emph{Global optimization with non-convex constraints. Sequential and parallel algorithms}. Kluwer Academic Publishers, Dordrecht.

\bibitem{Strongin18} Barkalov K., Strongin R. (2018) Solving a set of global optimization problems by the parallel technique with uniform convergence. \emph{J. Glob. Optim.}, Vol.~71(1), pp.~21--36.

\bibitem{Sergeyev16} Sergeyev Ya.D., Mukhametzhanov M.S., Kvasov D.E., Lera D. (2016) Derivative-free local tuning and local improvement techniques embedded in the univariate global optimization. \emph{J. Optim. Theory Appl.}, Vol.~171(1), pp.~186--208.

\bibitem{Sergeyev10} Lera D., Sergeyev Ya.D. (2010)  (2010) An information global minimization algorithm using the local improvement technique. \emph{J. Glob. Optim.}, Vol.~48(1), pp.~99--112.

\end{thebibliography}

\end{document}