### 2020-08-13 2020/OPTIMA/Answers.tex
Draft for the answers to the reviewers

%% Based on a TeXnicCenter-Template by Gyorgy SZEIDL.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------
%
\documentclass{article}%
%
\usepackage{amsmath}%
\usepackage{amsfonts}%
\usepackage{amssymb}%
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{ulem}
\usepackage{cancel}	
\usepackage[utf8]{inputenc}
\usepackage[english, russian]{babel} % Русские и английские переносы

%-------------------------------------------

\begin{document}

\section*{Answers to the reviewers' comments for the paper ``Parallel global optimization algorithm with uniform convergence for solving a set of constrained global optimization problems''}

We thank very much all the reviewers for their useful comments and suggestions.
Please find below our corrections that we suggest to clarify the issues.

\phantom{f}

\subsection*{Answers to Reviewer 1}

\subsubsection*{Major comments}

\begin{enumerate}
\item 
First of all, this is the Nizhny Novgorod version of the English demonstrated by the authors. ... I recommend that the authors should seek professional help from experienced translators of mathematical texts.

\textbf{Answer}. 
%The paper has been corrected by an English-speaking person.

\item The principal objective of the paper, which is to solve the number (set) of nonconvex problems, is not presented conveniently in the manuscript. There is no mathematical formulation of the problem, say, in the form of an optimization problem, meanwhile the authors tell us about “the optimal way” of solving such a problem, without substantiating this view.

\textbf{Answer}.

\item As well-known now, the solution of (1-3)-dimensional math programming problems, even nonconvex ones, has no considerable difficulties (say, with the help of B\&B methods) and, hence, it is not an attractive value for estimation of the power and effectiveness of algorithms.

\textbf{Answer}.

\item There is no comparison of the computational results with popular approaches in the Global Optimization, such as Random Searches, Genetics Algorithms (for example), and other bioinitiated methods.
Additionally, we recommend that the authors pay attention to the DCA methods (by P.D.Tao and L.Th.An) which are able of producing a stationary (KKT-) vector and even a global solution (sometimes) to DC problems.

\textbf{Answer}.

\end{enumerate}

\subsubsection*{Minor comments}

\begin{enumerate}
\item p. 3. Apparently, $z_k\neq g_s(x_k ) =:z_{ks}$.

\textbf{Answer}.

\item  “To account for the latter” change for “On account of”.

\textbf{Answer}.

\item  p. 4. To put more clear the definition (DF) of $Q_i$ (poor grammar usage).

\textbf{Answer}.

\item  Delete “At that” (???) globally.

\textbf{Answer}.

\item “At the point” put instead of “in the point”.

\textbf{Answer}.

\item  Specify the DF of the trial (what is it?) at the point $x$.

\textbf{Answer}.

\item  Where is a DF of $\nu(x)$?

\textbf{Answer}.

\item  Is it relevant to use at Step 1 “juxtapose”? What does it mean?

\textbf{Answer}.

\item  It seems that in Theorem 1 it is sufficient to assume that $D\neq\emptyset$ (I-st theorem of Weierstrass).

\textbf{Answer}.

\item  Delete “Beginning with some step”, put simply “IF …”

\textbf{Answer}.

\item  Change “admissible” for “feasible”.

\textbf{Answer}.

\item  The equalities (12) are wrong.

\textbf{Answer}.

\item  Delete “where $y_*$ is from (3)”.

\textbf{Answer}.

\item  Completely reformulate Theorem 2.

\textbf{Answer}.

\item  Before Table 1 the numbers (?) $p, S_i, S_t$ are not correctly defined.

\textbf{Answer}.

\item  Delete two last propositions before Table 1.

\textbf{Answer}.

\item  Completely rewrite the conclusion. As it is, it is not written in English.

\textbf{Answer}.

\item  The English and the terminology must be similar to [6] and [16].

\textbf{Answer}.

\end{enumerate}

\subsection*{Answers to Reviewer 2}

\textbf{Comment}. Is it possible to find different solutions by the suggested technique, please, give some comments.

\textbf{Answer}.

\subsection*{Answers to Reviewer 3}

\textbf{Comment}. It's not clear why solving a series of global optimization problems in parallel is an important problem. From my experience such problem statement sounds artificial.

\textbf{Answer}.

\end{document}