### 2020-09-29 2020/OPTIMA/slides/slides.tex
OPTIMA: fixes, update slides

\documentclass[aspectratio=1610]{beamer}

\usetheme{unnslides}
\usefonttheme{professionalfonts}

\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\inputencoding{utf8}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{cmbright}
\usepackage{fontspec}
\usepackage{unicode-math}
\usepackage{amsfonts}
\usepackage{subfig}
\usepackage{tikz}

\captionsetup[subfigure]{labelformat=empty}
\captionsetup[figure]{labelformat=empty}

\setmainfont{CMU Sans Serif}
\setromanfont{CMU Sans Serif}
\setsansfont{CMU Sans Serif}

\setlength{\tabcolsep}{1pt}

\usepackage{polyglossia}
\setmainlanguage{russian}
%\setbeamertemplate{itemize item}{\color{black}$\blacktriangleright$}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\re}{Re}

\graphicspath{ {../pic/}{img/} }

%set pages numeration
\newcommand\numbered{\setbeamertemplate{footline}{%
  \vspace{-10em}
   \raisebox{5pt}{\makebox[\paperwidth]{%
     \hfill\makebox[10pt]{%
       \usebeamerfont{footline}\usebeamercolor[fg]{footline}
       \insertframenumber}}}}}
\newcommand\unnumbered{\setbeamertemplate{footline}{}}

\title{Parallel global optimization algorithm with uniform convergence
for solving a set of constrained global optimization problems}
\author{\underline{\textbf{V.V.~Sovrasov}} \and \textbf{K.A.~Barkalov}}
\institute{Nizhny Novgorod State University}
\date{}

\begin{document}
\numbered
{
\unnumbered
\begin{frame}[noframenumbering,plain]
\titlepage
\end{frame}
}

\begin{frame}
  \frametitle{Problem statement}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{displaymath}
        \begin{array}{cr}\\
          \varphi(y^*)=\min\{\varphi(y):y\in G\},\:G=D \cap \{y:\\
           g_j(y)\leqslant 0, 1\leqslant j\leqslant m\},
        \end{array}
      \end{displaymath}
      where \(D\) is a hyperrectangle,
      \(\varphi(y),\:g_j(y)\) -- multi-extremal functions, satisfying the Lipschitz condition:
      \begin{displaymath}
        \begin{array}{cr}\\
          |f(y_1)-f(y_2)|\leqslant L\Vert y_1-y_2\Vert,\\y_1,y_2\in \mathbb{R}^N:a_i\leqslant y_i\leqslant{b_i},
        \end{array}
      \end{displaymath}
      where \(L>0\) is the Lipschitz constant and \(||\cdot||\) denotes the \(l_2\)-norm in \(\mathbb{R}^N\).
    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{img/gkls.png}}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Problem statement}
  Let's consider a set of \(q\) global optimization problems with nonlinear constraints:
  \begin{equation}
    \label{eq:many_problems}
    \min\left\{\varphi_1(y), y\in G_1 \right\}, \min\left\{\varphi_2(y), y\in G_2\right\},...,
  \min\left\{\varphi_q(y), y\in G_q\right\}.
  \end{equation}
  This work aims to formulate a method that will ensure the uniform convergence to the
  solutions of all problems in the series:
  \begin{equation}
    \exists \varepsilon > 0: \forall s>1, \forall i,j\in\{1,\dots,q\}\;
      \frac{\Vert \overline{y_i}^*(s) - y_i^*\Vert_\infty}{\Vert \overline{y_j}^*(s) - y_j^*\Vert_	\infty} \leqslant \varepsilon,
  \end{equation}
  where \(s\) is the number of step of an optimization method, \(\overline{y_i}^*(s)\) is the best iteration point
  in problem \(i\) from the set (\ref{eq:many_problems}) at the step \(s\).

\end{frame}

\begin{frame}
  \frametitle{Problem statement}
  Such sets of problems can arise in the following cases:
  \begin{itemize}
    \item continuous global optimization problem with a discrete parameter which takes a narrow range of values;
    \item a set of problems, obtained after scalarization of a multi-objective problem.
  \end{itemize}
\textbf{Possible approaches}:
  \begin{itemize}
    \item solve each problem independently;
    \item develop an optimiaztion method, which solves all the problems "simultaneously", at each moment of time
    prioritizing one of the problems.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Dimension reduction}
  \begin{center}
  Peano-type curve \(y(x)\) allows to reduce the dimension of the original problem:
  \begin{gather}
    D_e=\lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
    \min\{\varphi(y): y\in D\}=\min\{\varphi(y(x)): x\in [0,1]\} \nonumber
  \end{gather}
  \(y(x)\) is non-smooth function which continuously maps the segment \([0,1]\) to the hypercube \(D_e\).
  \begin{figure}[ht]
    \includegraphics[width=.35\textwidth]{peano3d.png}
    \includegraphics[width=.35\textwidth]{peano2d.png}
  \end{figure}
\end{center}
\end{frame}

\begin{frame}
  \frametitle{Univariate Algorithm of Global Search}
  Optimization method generates search sequence \(\{x_k:x_k\in[a,b]\}\) and consists of the following steps:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
    \item[Step 2.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
    \item[Step 3.] Choose an interval with the greatest characteristic, \((x_{t-1}, x_{t})\), and
    make a trial (compute the constraints and objective) at the point \(x^{k+1}\), chosen according with decision rule \(d\):
    \begin{displaymath}
      x^{k+1}=d(t)\in (x_{t-1}, x_{t})
    \end{displaymath}
    \item[Step 4.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
  \end{enumerate}
  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
\end{frame}

\begin{frame}
  \frametitle{Algorithm for solving a set of problems}
  The method relies on the previously introduces AGS and consists of the following steps:
  \begin{itemize}
    \item Create \(q\) copies of AGS.
    \item Run \(q\) synchronously working copies of AGS, but instead of performing Step 3, stop all
    the methods and select an interval with the highest characteristic among intervals from all the methods.
    \item If the interval with the highest characteristic belongs to problem i \(i\), execute Step 3 in the \(i^{th}\) copy of AGS.
    Other copies skip Step 3.
    \item To implement parallel processing, method picks \(p\) intervals at the previous two steps
    and performs \(p\) trials simultaneously.
  \end{itemize}
  The described method relies to \textbf{comparability of characteristics in different problems}.
\end{frame}

\begin{frame}
  \frametitle{Index method of handling inequality constraints}
  Instead of the original problems with functional constraints the index scheme considers the following unconstrained problem:
  \begin{displaymath}
    \begin{array}{lr}
      \psi (x^{*})=\min_{x\in [0;1]}\psi (x), \\
      \psi (x)={\begin{cases}g_{\nu }(x)/H_{\nu }&\nu <M\\(g_{M}(x)-g_{M}^{*})/H_{M}&\nu =M\end{cases}},
    \end{array}
  \end{displaymath}
  where if \(\nu=m+1\;,g_\nu(x)=\varphi(x)\).

  Characteristics are defined as follows:
  \begin{displaymath}
    R(i)={\begin{cases}\Delta _{i}+{\frac {(z_{i}-z_{i-1})^{2}}{(r_{\nu }\mu _{\nu })^{2}\Delta _{i}}}-2{\frac {z_{i}+z_{i-1}-2z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i})=\nu (x_{i-1})\\2\Delta _{i}-4{\frac {z_{i-1}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i-1})>\nu (x_{i})\\2\Delta _{i}-4{\frac {z_{i}-z_{\nu }^{*}}{r_{\nu }\mu _{\nu }}}&\nu =\nu (x_{i})>\nu (x_{i-1})\end{cases}}
  \end{displaymath}
  where \(z_i=\psi(x_i)\).
\end{frame}

\begin{frame}
  \frametitle{Index method of handling inequality constraints}
  \begin{enumerate}
    \item \(D\ne\emptyset\), the problem (\ref{eq:constrained_problem}) has a solution.
    \item Functions \(g_j(y)\leqslant 0, 1\leqslant j\leqslant m + 1\), are Lipschitzian with
respective constants \(L_i\) over the domain \(D\) (here \(g_{m+1}(y)=\varphi(y)\)).
    \item If \(k\) from (\ref{eq:points}) is sufficiently large,
    the values \(\mu_\nu\) from (\ref{step2}) satisfy the inequalities
    \begin{equation}
      r_\nu\mu_\nu > 2^{3-1/N}L_\nu \sqrt{N+3},\: 1\leqslant \nu \leqslant m + 1.
    \end{equation}
  \end{enumerate}
  \begin{theorem} (On convergence conditions of MIAGS) Let for each problem \(i,\:1\leqslant i\leqslant q\) from (\ref{eq:many_problems})
    the conditions 1-3 of the Theorem \ref{th:single_conv} are true, i.e. each problem can be solved by IAGS.
    Then solving all of the \(q\) problems by MIAGS will
    generate \(q\) infinite sequences \(\{y^k_i\},\:1\leqslant i\leqslant q\), such that
    \begin{displaymath}
      \varphi_i(\overline{y_i})=\inf\{ \varphi(y^k_i): g^i_j(y^k_i)\leqslant 0,1\leqslant j\leqslant m_i, k=1,2,\dots\}=\varphi_i(y^*_i).
    \end{displaymath}
  \end{theorem}
\end{frame}

\begin{frame}
  \frametitle{An example of solving a multi-objective problem}
  Consider the following problem:
  \begin{displaymath}
    \begin{array}{l}
        Minimize \left \{
        \begin{array}{l}
          f_1(y) = 4 y_1^2 + 4 y_2^2 \\
          f_2(y) = (y_1-5)^2 + (y_2-5)^2 \\
        \end{array}
        \right .
        y_1\in [-1;2],y_2\in [-2;1]
        \\s.t.
        \\
        \left \{
        \begin{array}{l}
          g_1(y) = (y_1 - 5)^2 + y_2^2 - 25 \leqslant 0 \\
          g_2(y) = -(y_1 - 8)^2 - (y_2 + 3)^2 + 7.7 \leqslant 0\\
        \end{array}
        \right .
    \end{array}
  \end{displaymath}
  After scalarization the problem has the following form:
  \begin{displaymath}
    \varphi(y^*(\lambda_1,\lambda_2))=\min_{y\in D}\max\{\lambda_1 f_1(y), \lambda_2 f_2(y)\};\lambda_1,\lambda_2\in[0,1],\: \lambda_1+\lambda_2=1
\end{displaymath}
To build the Pareto set numerically, consider
100 pairs of the variables \((\lambda_1,\lambda_2)\) such that
\(\lambda_1^i=i h,\: \lambda_2^i=1-\lambda_1^i,\: h=10^{-2},i=\overline{1, 100}\).
\end{frame}

\begin{frame}
  \frametitle{An example of solving a multi-objective problem}
  \vspace*{-1.0cm}
  \begin{displaymath}
    \begin{array}{cr}\\
      SP(S)=\sqrt{\frac{1}{|S|-1} \sum_{i=1}^{|S|} (\overline{d}-d_i)^2},\:\overline{d}=mean\{d_i\}, \\
      d_i=\min_{s_i,s_j\in S:s_i\ne s_j}||F(s_i)-F(s_j)||_1,\: F=(f_1,f_2)
    \end{array}
  \end{displaymath}

  \begin{figure}[ht]
      \centering
      \vspace*{-0.5cm}
      \subfloat[IAGS, \(SP_{single}=0.984\)]{
      \label{fig:mco_pareto_1} {\includegraphics[width=.5\textwidth]{single_mco.pdf}}}
      \subfloat[MIAGS, \(SP_{multi}=0.749\)]{
      \label{fig:mco_pareto_2} {\includegraphics[width=.5\textwidth]{multi_mco.pdf}}}
      \caption{Numerical estimations of the Pareto set, obtained after 2500 trails}
      \label{fig:mco_pareto}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Generated test problems}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      Sets of test problems are obtained by applying the GCGen system to
      functions produced by generators of unconstrained test problems.
      The main features of the generated sets:
      \begin{itemize}
        \item \(q\)=100;
        \item Dimension 2, 3 or 4;
        \item Base fuctions GKLS or \(F_{GR}\), or a combination of them;
        \item Number of nonlinear constraints: 2.
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{4.png}}
    \end{column}
  \end{columns}
  \footnotesize{GCGen is available here \textit{https://github.com/UNN-ITMM-Software/GCGen}}
\end{frame}

\begin{frame}
  \frametitle{Hardware and software environment}
  \begin{center}
    The parallel method was implemented on C++ language with the use of OpenMP
    technology for organizing parallelism on shared memory.

    All the numerical experiments were carried out on a computer with the following
    configuration: Intel Core i7-7800X (6 cores) CPU, 64GB RAM, Unubtu 16.04 ОS, GCC 5.5 compiler.
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Result of solving sets of synthetic problems}
  \begin{figure}[ht]
    \vspace*{-0.5cm}
      \centering
      \subfloat[\(D_{max}\)]{
      \label{fig:max_dev} {\includegraphics[width=.5\textwidth]{mixed_2d_max.pdf}}}
      \subfloat[\(D_{avg}\)]{
      \label{fig:avg_dev} {\includegraphics[width=.5\textwidth]{mixed_2d_avg.pdf}}}
      \caption{Dynamic of the average and maximal errors, during solving of problems
      generated on different types of base functions: GKLS и \(F_{GR}\)}
      \label{fig:devs_mixed}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Result of solving sets of synthetic problems}
  \begin{table}
    \centering
    \begin{tabular}{c|c|c|c|c|c}
      %\cline{1-8}\noalign{\smallskip}
      Problems class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \hline
      GKLS \& \(F_{GR}\)-based \
        & 1 & 51434 & 90.20 & -    & - \\
        & 2 & 25698 & 56.96 & 2.00 & 1.58 \\
        & 4 & 13015 & 36.67 & 3.95 & 2.46 \\
        & 6 & 8332  & 26.85 & 6.17 & 3.36 \\
      \hline
      GKLS-based 2d \
        & 1 & 59066 & 97.53 & -    & - \\
        & 2 & 29060 & 60.56 & 2.04 & 1.61 \\
        & 4 & 14266 & 38.92 & 4.14 & 2.51 \\
        & 6 & 9436  & 29.53 & 6.26 & 3.30 \\
      \hline
    \end{tabular}
  \end{table}
\end{frame}
\begin{frame}
  \frametitle{Result of solving sets of synthetic problems}
  \begin{table}
    \centering
    \begin{tabular}{c|c|c|c|c|c}
      %\cline{1-8}\noalign{\smallskip}
      Problems class & \textit{p} & Number of iterations & Time, s & \(S_i\) & \(S_t\)   \\
      %s\noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \hline
      GKLS-based 3d \
        & 1 & 782544 & 1117.55 & -    & - \\
        & 2 & 397565 & 752.92  & 1.97 & 1.48 \\
        & 4 & 208073 & 526.67  & 3.76 & 2.12 \\
        & 6 & 142089 & 445.45  & 5.50 & 2.51 \\
      \hline
      GKLS-based 4d \
        & 1 & 14021720 & 15806.6 & -    & - \\
        & 2 & 6313070 & 7254.85  & 2.22 & 2.18 \\
        & 4 & 3479344 & 4932.55  & 4.03 & 3.20 \\
        & 6 & 2783339 & 3955.38  & 5.04 & 3.99 \\
      \hline
    \end{tabular}
  \end{table}
\end{frame}
\begin{frame}
  \frametitle{Conclusions}
    \begin{itemize}
      \item implemented support of nonlinear functional constraints in the method
      for solving a set of global optimization problems;
      \item convergence of the implemented method was theoretically proven;
      \item conducted numerical experiments on sets, demonstrating uniform convergence (in the mentioned sense) and
      parallel speedup;
      \item demonstrated efficiency of the implemented method in case of solving a multi-objective problem.
    \end{itemize}
\end{frame}

{
\unnumbered
\begin{frame}{{}}
  \frametitle{Q\&A}
  \begin{center}
    %\Large{Контакты:}
\vspace{0.5cm}

    sovrasov.vlad@gmail.com

    https://github.com/sovrasov
  \end{center}
\end{frame}
}
\end{document}