### 2020-12-10 2021/PСT_GergelKozinov/2021_PCT.tex
PCT 2021/PСT_GergelKozinov 1. Оформлен список литературы

\documentclass{svproc}

\usepackage{url}
\usepackage{hyperref}

\usepackage{adjustbox}
\usepackage{amsmath}

\begin{document}

\title{Parallel Computations for Solving Multicriteria Mixed-Integer Optimization Problems}

\author{Victor Gergel \and Evgeniy Kozinov}

\institute{Institute of Informational Technologies, Mathematics and Mechanics, Mathematical Center, Lobachevsky State University of Nizhny Novgorod, Nizhni Novgorod, Russia \\
\email{gergel@unn.ru, evgeny.kozinov@itmm.unn.ru}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
abstract.
\keywords{Global optimization $\cdot$ dimensionality reduction $\cdot$ search information $\cdot$ parallel methods of global search $\cdot$ computational complexity}
\end{abstract}

\section{Introduction}\label{sec:1}

\cite{c1,c2,c3,c4,c5,c6,c7,c8,c9,c10}
\cite{c11,c12,c13,c14,c15,c16,c17,c18,c19,c20}
\cite{c21,c22,c23,c24,c25,c26,c27,c28,c29,c30}
\cite{c31,c32,c33,c34,c35,c36,c37,c38,c39,c40}
\cite{c41,c42,c43,c44,c45}

\section*{Acknowledgements} 
This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''

\bibliographystyle{spmpsci}
%\bibliographystyle{unsrt}
\bibliography{mybibfile}

\end{document}